### 2021-05-30 group-a/group-a-analysis.Rmd
add projects

---
title: "Group A"
output: html_notebook
---

# Setup

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(jsonlite)
library(ggplot2)
library(readr)
library(afex)
```

# Import Data

Read JSON files and bind together into a tibble.

```{r}
data.files <- list.files('data', full.names = TRUE, pattern=".json")
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  return(data.table)
})
all.data <- bind_rows(data.tables)
```

Get a list of subject IDs.

```{r}
subjects <- unique(all.data$subject_id)
```

# Extract Relevant Data

Get data from critical trials, select relevant columns, unnest/unpack webgazer data to get a tidy tibble,
and add gaze measurements as a percentage.

```{r}
task.data <- all.data %>%
  filter(verb_type %in% c("restricting", "non-restricting")) %>%
  select(subject_id, stimulus, response, verb_type, webgazer_data, webgazer_targets) %>%
  tidyr::unpack(webgazer_targets) %>%
  tidyr::unpack(`#scenes`) %>%
  select(-x,-y) %>%
  unnest(webgazer_data) %>%
  mutate(gaze.x.percent = (x - left) / width * 100) %>%
  mutate(gaze.y.percent = (y - top) / height * 100)
```

# Add Stimulus Info

Merge in info on visual ROIs and audio onsets.

```{r message=FALSE, warning=FALSE}
ref.image.height <- 1500 # height of images that were measured for ROIs
ref.image.width <- 1875 # width of images that were measured for ROIs
padding <- 20 # padding in pixels around the ROIs.

image.data <- read_csv('info/image_measurements.csv') %>%
  mutate(roi_top = (top_object_distance - padding) / ref.image.height * 100,
         roi_bottom = (bottom_object_distance + padding) / ref.image.height * 100,
         roi_left = (left_object_distance - padding) / ref.image.width * 100,
         roi_right = (right_object_distance + padding) / ref.image.width * 100) %>%
  mutate(object_class = if_else(object_type %in% c('subject', 'target.object'), object_type, 'distractor')) %>%
  select(image_name, object_type, object_class, roi_top, roi_bottom, roi_left, roi_right)

audio.data <- read_csv('info/audio_timestamps.csv') %>% filter(!is.na(stimulus))

critical.trials <- task.data %>%
  left_join(audio.data) %>%
  left_join(image.data) 
```

# Compute Gaze in ROI

Define a function for determining if `x,y` value falls in box.

```{r}
in.box <- function(x, y, left, right, top, bottom, padding){
  is.in.the.box <- x >= left - padding & x <= right + padding & y >= top - padding & y <= bottom + padding
  return(is.in.the.box)
}
```

Add a column with boolean value for whether gaze was in the ROI.

```{r}
critical.trials <- critical.trials %>%
  mutate(is.fixating = in.box(gaze.x.percent, gaze.y.percent, roi_left, roi_right, roi_top, roi_bottom, 0))
```

Check the proportion of samples that were in at least one ROI for subjects.

```{r}
prop.in.roi.by.subject <- critical.trials %>%
  group_by(subject_id, t, image_name) %>%
  summarize(fixation.in.roi = any(is.fixating)) %>%
  group_by(subject_id) %>%
  summarize(prop.of.samples.in.roi = mean(fixation.in.roi))
```
Visualize proportion of samples in ROI by subject.

```{r}
hist(prop.in.roi.by.subject$prop.of.samples.in.roi, breaks=20)
```
Exclude subjects below a threshold?

```{r}
bad.subjects <- prop.in.roi.by.subject %>%
  filter(prop.of.samples.in.roi == 0) %>%
  pull(subject_id)

# to match laila's analysis
bad.subjects <- c("5f3a76043bb11c6786e731bf", "5f9c6b2def8bdf1102de5338", "607197e359bae0449ac9d4bc", "60142d5718e5943b09fd9298")

critical.trials.filtered <- critical.trials %>%
  filter(!subject_id %in% bad.subjects)
```

# Calculating cumulative fixation probability

We start by calculating cumulative fixation probability on the objects after the verb onset. (Any fixations prior to the verb onset are not relevant to the core research questions.)

First, add a column for time relative to verb onset, bin this to 50ms chunks, and filter the data to the 2,000 ms window after verb onset.

```{r}
cumulative.fixation.data <- critical.trials.filtered %>%
  mutate(t.relative.to.verb.onset = t - (verb_onset*1000)) %>%
  filter(t.relative.to.verb.onset >= 0 & t.relative.to.verb.onset <2000) %>%
  mutate(t.w = floor(t.relative.to.verb.onset/50)*50)
```

Now check every time window within a trial to account for sparse samples. This calculating is time consuming, so we cache the result in a CSV file, and load from the CSV if it already exists.

```{r}
cumulative.fixation.calculation <- function(df){
  time <- seq(0,1950,50)
  fixations <- logical(length(time))
  hit.flag <- FALSE
  for(i in 1:length(fixations)){
    if(hit.flag){
      fixations[i] <- TRUE
    } else {
      val <- df %>% filter(t.w <= time[i]) %>% pull(is.fixating) %>% any()
      if(val == TRUE) {
        hit.flag <- TRUE
      }
      fixations[i] <- val
    }
  }
  out <- tibble(t.window = time, has.fixated = fixations)
  return(out)
}

if(file.exists("data/generated/cumulative_fixation_data.csv")){
  cumulative.fixation.windows.data <- read_csv("data/generated/cumulative_fixation_data.csv")
} else {

  cumulative.fixation.windows.data <- cumulative.fixation.data %>%
    group_by(subject_id, verb_type, image_name, object_type, object_class) %>%
    summarize(cumulative.fixation.calculation(cur_data()))

  write_csv(cumulative.fixation.windows.data, path="data/generated/cumulative_fixation_data.csv")
}
```

Then group by object class to calculate fixation probability for each kind of object, normalizing by the number of objects of that type.

```{r}
cumulative.fixation.windows.grouped.data <- cumulative.fixation.windows.data %>%
  filter(!subject_id %in% bad.subjects) %>%
  group_by(subject_id, verb_type, image_name, object_class, t.window) %>%
  summarize(cumulative.fixation.p = mean(has.fixated)) 
```

Then collapse over trials to get a cumulative probability for the whole window for each subject

```{r}
cumulative.fixation.windows.grouped.average.data <- cumulative.fixation.windows.grouped.data %>%
  group_by(subject_id, verb_type, object_class, t.window) %>%
  summarize(cumulative.fixation.p = mean(cumulative.fixation.p))
```

Collapse across subjects to generate the equivalent of the figure from Altmann & Kamide.

```{r}
cumulative.fixation.data.summary <- cumulative.fixation.windows.grouped.average.data %>%
  filter(object_class %in% c('target.object', 'distractor')) %>%
  group_by(t.window, verb_type, object_class) %>%
  summarize(cumulative.fixation.m = mean(cumulative.fixation.p))
```

```{r}
ggplot(cumulative.fixation.data.summary, aes(x=t.window, y=cumulative.fixation.m, fill=object_class, shape=verb_type, group=interaction(verb_type, object_class)))+
  scale_fill_grey(start=1.0, end=0.0)+
  scale_shape_manual(values=c(22,21))+
  geom_line()+
  geom_point(size=2)+
  theme_bw()+
  theme(panel.grid=element_blank())+
  scale_y_continuous(limits=c(0,1))+
  scale_x_continuous(expand=c(0.01,0.01))+
  labs(x="Time from Verb Onset", y="Probability", shape="Verb Type", fill="Object Type")

```

# Question: Are there more pre-noun first looks to the target object than to the distractors?

First, grab the noun onset times to merge into our cumulative fixation data.

```{r}
noun.onset.times <- critical.trials.filtered %>%
  group_by(image_name, verb_type) %>%
  slice(1) %>%
  ungroup() %>%
  select(image_name, verb_type, delay_bw_onset_of_verb_and_onset_of_target_noun) %>%
  mutate(relative.noun.onset = delay_bw_onset_of_verb_and_onset_of_target_noun * 1000) %>%
  select(-delay_bw_onset_of_verb_and_onset_of_target_noun)

trial.level.cumulative.fixation.data <- cumulative.fixation.windows.grouped.data %>%
  left_join(noun.onset.times, by=c("image_name", "verb_type"))
```

Then, grab the last time window before the noun onset for each trial. This is the critical data for this test.

```{r}
pre.noun.first.looks.data <- trial.level.cumulative.fixation.data %>%
  group_by(subject_id, image_name, verb_type) %>%
  filter(t.window < relative.noun.onset) %>%
  filter(t.window == max(t.window)) %>%
  filter(object_class %in% c("target.object", "distractor"))
```

Fit the model, with no covariance between random effects because the model cannot converge with full covariance matrix.

```{r}
model <- afex::lmer_alt(cumulative.fixation.p ~ object_class * verb_type + (object_class * verb_type || image_name) + (object_class * verb_type || subject_id) , data=pre.noun.first.looks.data)
summary(model)
```

# Question: Are there more pre-verb-offset first looks to the target object than to the distractors?

Verb offset times

```{r}
verb.offset.times <- critical.trials.filtered %>%
  group_by(image_name, verb_type) %>%
  slice(1) %>%
  ungroup() %>%
  select(image_name, verb_type, duration_of_verb) %>%
  mutate(relative.verb.offset = duration_of_verb * 1000) %>%
  select(-duration_of_verb)

trial.level.cumulative.fixation.data.verb.offset <- cumulative.fixation.windows.grouped.data %>%
  left_join(verb.offset.times, by=c("image_name", "verb_type"))
```

```{r}
pre.verb.offset.first.looks.data <- trial.level.cumulative.fixation.data.verb.offset %>%
  group_by(subject_id, image_name, verb_type) %>%
  filter(t.window < relative.verb.offset) %>%
  filter(t.window == max(t.window)) %>%
  filter(object_class %in% c("target.object", "distractor"))
```

```{r}
model <- lmer_alt(cumulative.fixation.p ~ object_class * verb_type + (object_class * verb_type || image_name) + (object_class * verb_type || subject_id) , data=pre.verb.offset.first.looks.data)
summary(model)
```

# Do people look at the target faster after a restricting verb

Relative to verb onset, when is the first fixation on the target?

```{r}
first.fixation.after.verb.onset <- critical.trials.filtered %>%
  group_by(subject_id, image_name) %>%
  filter(t >= verb_onset*1000) %>%
  filter(object_type == "target.object") %>%
  filter(is.fixating == TRUE) %>%
  filter(t == min(t)) %>%
  mutate(relative.t = t - (postverbal_noun_onset*1000))
```

```{r}
first.fixation.after.verb.onset.analysis.data <- first.fixation.after.verb.onset %>%
  ungroup() %>%
  select(subject_id, image_name, verb_type, relative.t)
```

```{r}
first.fixation.after.verb.onset.analysis.data %>%
  group_by(verb_type, subject_id) %>%
  summarize(M=mean(relative.t)) %>%
  ggplot(aes(x=verb_type, y=M, group=subject_id)) +
  geom_point()+
  geom_line()
```

```{r}
model <- lmer_alt(relative.t ~ verb_type + (verb_type || subject_id) + (verb_type || image_name), data=first.fixation.after.verb.onset.analysis.data)

summary(model)
```

```{r}
first.fixation.after.verb.onset.subject.summary <- first.fixation.after.verb.onset %>%
  group_by(subject_id, verb_type) %>%
  summarize(m.t = mean(relative.t), ct = n())

first.fixation.after.verb.onset.summary <- first.fixation.after.verb.onset.subject.summary %>%
  group_by(verb_type) %>%
  summarize(m=mean(m.t), sd=sd(m.t), se=sd/sqrt(n()))
```

```{r}
ggplot(first.fixation.after.verb.onset.summary, aes(x=verb_type, y=m, ymin=m-se,ymax=m+se))+
  geom_col()+
  geom_errorbar(width=0.2)
```