### 2021-05-30 group-b/group-b-analysis.Rmd

---
title: "Group B Analysis"
output: html_notebook
---

```{r message=FALSE}
library(jsonlite)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ez)
```

# Import Data

Read JSON files and bind together into a tibble.

```{r}
data.files <- list.files('data', full.names = TRUE, pattern=".json")
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  return(data.table)
})
all.data <- bind_rows(data.tables)
```

# Extract relevant data

```{r}
task.data <- all.data %>%
  filter(!is.na(relation)) %>%
  select(subject_id, trial_index, stimulus, relation, quadrant, rt, correct_response, response, webgazer_data, webgazer_targets) %>%
  unpack(webgazer_targets) %>%
  unpack(`#screen`) %>%
  select(-top,-left,-x,-y,-bottom,-right)
```

Need to add condition information to the `task.data`

```{r}
condition.subject.info <- all.data %>%
  filter(!is.na(condition)) %>%
  select(subject_id, condition) %>%
  group_by(subject_id) %>%
  mutate(block = c(1,2)) %>%
  ungroup()

task.data <- task.data %>%
  group_by(subject_id) %>%
  mutate(block = c(rep(1,n()/2),rep(2,n()/2))) %>%
  left_join(condition.subject.info, by=c("subject_id", "block"))
```

# Question: Do people look more at the critical quadrant during free viewing?

Grab the relevant data, unnesting webgazer observations

```{r}
free.view.data <- task.data %>%
  filter(condition == "free") %>%
  unnest(webgazer_data)
```

Add quadrant information to each observation

```{r}
free.view.data <- free.view.data %>%
  mutate(x.percent = x / width * 100, y.percent = y / height * 100) %>%
  mutate(view_quadrant = case_when(
    x.percent <= 50 & y.percent <= 50 ~ 'top.left',
    x.percent > 50 & y.percent <= 50 ~ 'top.right',
    x.percent <= 50 & y.percent > 50 ~ 'bottom.left',
    x.percent > 50 & y.percent > 50 ~ 'bottom.right'
  )) %>%
  mutate(normalized_quadrant = case_when(
    quadrant == 1 & view_quadrant == 'top.left' ~ 'critical',
    quadrant == 1 & view_quadrant == 'top.right' ~ 'first',
    quadrant == 1 & view_quadrant == 'bottom.right' ~ 'second',
    quadrant == 1 & view_quadrant == 'bottom.left' ~ 'third',

    quadrant == 2 & view_quadrant == 'top.left' ~ 'third',
    quadrant == 2 & view_quadrant == 'top.right' ~ 'critical',
    quadrant == 2 & view_quadrant == 'bottom.right' ~ 'first',
    quadrant == 2 & view_quadrant == 'bottom.left' ~ 'second',

    quadrant == 3 & view_quadrant == 'top.left' ~ 'first',
    quadrant == 3 & view_quadrant == 'top.right' ~ 'second',
    quadrant == 3 & view_quadrant == 'bottom.right' ~ 'third',
    quadrant == 3 & view_quadrant == 'bottom.left' ~ 'critical',

    quadrant == 4 & view_quadrant == 'top.left' ~ 'second',
    quadrant == 4 & view_quadrant == 'top.right' ~ 'third',
    quadrant == 4 & view_quadrant == 'bottom.right' ~ 'critical',
    quadrant == 4 & view_quadrant == 'bottom.left' ~ 'first'
  ))
```

Summarize proportion of views in each quadrant for each trial
```{r}
free.view.summary.trial.data <- free.view.data %>%
  group_by(subject_id, trial_index, relation) %>%
  summarize(critical = sum(normalized_quadrant == 'critical')/n(),
            first = sum(normalized_quadrant == 'first')/n(),
            second = sum(normalized_quadrant == 'second')/n(),
            third = sum(normalized_quadrant == 'third')/n()) %>%
  pivot_longer(c("critical", "first", "second", "third"), names_to = "normalized_quadrant", values_to = "proportion")
```
Summarize by subject
```{r}
free.view.summary.subject.data <- free.view.summary.trial.data %>%
  group_by(subject_id,relation,normalized_quadrant) %>%
  summarize(proportion = mean(proportion))
```

Summarize by condition 
```{r}
free.view.summary.condition.data <- free.view.summary.subject.data %>%
  group_by(normalized_quadrant, relation) %>%
  summarize(M = mean(proportion), SE = sd(proportion) / sqrt(n()))
```
```{r}
ggplot(free.view.summary.condition.data, aes(x=normalized_quadrant, y=M, ymax=M+SE, ymin=M-SE, fill=relation)) +
  geom_col(position=position_dodge())+
  geom_errorbar(position=position_dodge(width=0.9), width=0.1)+
  theme_classic()
```

Goal: ANOVA with quadrant and statement type as factors, gaze proportion as DV

```{r}
free.view.anova <- ezANOVA(free.view.summary.subject.data, wid=subject_id, dv=proportion, within=c(relation,normalized_quadrant))
free.view.anova$ANOVA
```
# Behavioral Analysis

First, extract relevant data and compute `correct` column

```{r}
behavioral.data <- task.data %>%
  select(subject_id, trial_index, relation, rt, response, correct_response, condition) %>%
  mutate(correct = response == correct_response)
```

Goal: ANOVA with RT as DV, condition, statement as IVs

```{r}
rt.behavioral.subject.data <- behavioral.data %>%
  group_by(subject_id, relation, condition) %>%
  filter(correct == TRUE) %>%
  summarize(rt = mean(rt))
```

```{r}
rt.anova <- ezANOVA(rt.behavioral.subject.data, dv=rt, wid=subject_id, within = c(relation, condition))
rt.anova$ANOVA
```

```{r}
rt.summary.condition.data <- rt.behavioral.subject.data %>%
  group_by(relation, condition) %>%
  summarize(M=mean(rt), SE=sd(rt)/sqrt(n()))

ggplot(rt.summary.condition.data, aes(x=relation, fill=condition, color=condition, y=M, ymax=M+SE, ymin=M-SE, group=condition))+
  geom_point(size=5)+
  geom_line()+
  geom_errorbar(width=0.2)+
  scale_color_manual(values=c("orange2","royalblue4"))+
  labs(x="Statement Type", y="Accuracy", color=NULL)+
  theme_classic()
```

Goal: ANOVA with ACC as DV, condition, statement as IVs

*note that paper computer accuracy as hit rate - false alarm rate*
```{r}
acc.behavioral.subject.data <- behavioral.data %>%
  group_by(subject_id, relation, condition) %>%
  summarize(hit.rate = sum(correct == TRUE & response == 't') / sum(correct_response == 't'),
            fa.rate = sum(correct == FALSE & response == 't') / sum(correct_response == 'f')) %>%
  mutate(accuracy = hit.rate - fa.rate)
```

```{r}
acc.anova <- ezANOVA(acc.behavioral.subject.data, dv=accuracy, wid=subject_id, within = c(relation, condition))
acc.anova$ANOVA
```

Recreate figure from paper

```{r}
acc.summary.condition.data <- acc.behavioral.subject.data %>%
  group_by(relation, condition) %>%
  summarize(M=mean(accuracy), SE=sd(accuracy)/sqrt(n()))

ggplot(acc.summary.condition.data, aes(x=relation, color=condition, y=M, ymax=M+SE, ymin=M-SE, group=condition))+
  geom_point(size=5)+
  geom_line()+
  geom_errorbar(width=0.2)+
  scale_color_manual(values=c("orange2","royalblue4"))+
  coord_cartesian(ylim=c(0.3,0.8)) +
  labs(x="Statement Type", y="Accuracy", color=NULL)+
  theme_classic()
```