### 2021-11-11 conference/CDS-abstract-figure.Rmd
working on figure for CDS abstract

---
title: "CDS-abstract-figure"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(patchwork)
library(magick)
library(ggpubr)
```

```{r }

displayA<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.001.jpeg'))+coord_fixed() + ggtitle('A')
displayB<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.002.jpeg'))+coord_fixed() + ggtitle('B')
displayC<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.003.jpeg'))+coord_fixed() + ggtitle('C')
displayD<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.004.jpeg'))+coord_fixed() + ggtitle('D')
displayE<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.005.jpeg'))+coord_fixed() + ggtitle('E')

figA<-read_rds('../group-d/output/ETfig.rds')
figB<-read_rds('../group-d/output/ETfig.rds')
figC<-read_rds('../group-d/output/ETfig.rds')
figD<-read_rds('../group-d/output/ETfig.rds')
figE<-read_rds('../group-d/output/ETfig.rds')

```

```{r}
A <- displayA + figA 
B <- displayA + figB 
C <- displayA + figC 
D <- displayD + figD 
E <- displayA + figE

```

```{r}
(A | B)/(C | D ) / ( E | (grid::textGrob('Figure 1. ADD CAPTION HERE')) ) 
```

```{r}
ggsave('CDS-abstract-figure.pdf', scale = 1, width = 11, height = 7)

```