### 2021-11-13 conference/CDS-abstract-figure.Rmd
all figures added for CDS

---
title: "CDS-abstract-figure"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(patchwork)
library(magick)
library(ggpubr)
```

```{r }

displayA<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.001.jpeg'))+coord_fixed() + ggtitle('A')
displayB<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.002.jpeg'))+coord_fixed() + ggtitle('B')
displayC<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.003.jpeg'))+coord_fixed() + ggtitle('C')
displayD<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.004.jpeg'))+coord_fixed() + ggtitle('D')
displayE<-ggplot() + background_image(image_read('expt-example-displays/expt-example-displays.005.jpeg'))+coord_fixed() + ggtitle('E')

figA<-read_rds('../group-a/output/exp1_plot.rds')
figB<-read_rds('../group-b/output/ETfig.rds')
figC<-read_rds('../group-c/output/exp3_plot.rds')+scale_color_brewer(palette = "Set1")
figD<-read_rds('../group-d/output/ETfig.rds')
figE<-read_rds('../group-e/output/exp5_plot.rds')+scale_color_brewer(palette = "Set1")

figA.1<-ggplot(figA$data, 
               aes(x=t.window, y=cumulative.fixation.m, shape = object_class, color = verb_type))+
  geom_point(size=1)+
  geom_line()+
  scale_color_brewer(palette = "Set1", name = "Verb type")+
  scale_shape_manual(name = "Gaze location", values = c(1,4), labels = c("distractor", "target"))+
  theme_classic()+
  labs(x = "Time from Verb Onset", y="Cumulative Look Probability")
figA.1
```

```{r}
A <- displayA + figA.1 + plot_layout(widths = c(1.5,1)) & theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))
B <- displayB + figB + plot_layout(widths = c(1.5,1)) & theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))
C <- displayC + figC + plot_layout(widths = c(1.5,1)) & theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))
D <- displayD + figD + plot_layout(widths = c(1.5,1)) & theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))
E <- displayE + figE + plot_layout(widths = c(1.5,1)) & theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))

```

```{r}
(A | B)/(C | D ) / ( E | (grid::textGrob('Figure 1. ADD CAPTION HERE')) ) 
```

```{r}
ggsave('CDS-abstract-figure.pdf', width = 11.5, height = 7)

```