### 2022-01-09 group-d/E4-analysis-comparison-to-original.Rmd
added data from ryskin et al 2017 for comparison to group d

---
title: "E4-analysis-comparison-to-original"
author: "Rachel Ryskin"
output: pdf_document
---

```{r , include=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
```

```{r E4-load-web-data}
first.move = read_csv("output/E4_mouse_data.csv")
eyetracking.window.summary.by.trial = read_csv( "output/E4_eye-tracking_data.csv")
```

```{r load-data-from-original-study}

first.move.orig = read_tsv("original_study_data/Experiment1_clickData.txt")
eyetracking.window1.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_FirstTimeWindow.txt")
eyetracking.window2.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_SecondTimeWindow.txt")
eyetracking.window3.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_ThirdTimeWindow.txt")
```

```{r E4-summarizing-first-mouse-moves-for-plotting-orig}
# Summarize the data by subject, calculating proportion of trials the the first move was to the animal, instrument, or other.
first.move.orig.subject.summary <- first.move.orig %>%
  group_by(subject, condition) %>%
  summarize(prop.animal = mean(clickedTargetAnimalFirst),
            prop.instrument = mean(clickedTargetInstrumentFirst),
            prop.other = mean(clickedTargetAnimalFirst == 0 & clickedTargetInstrumentFirst == 0 )) %>%
  pivot_longer(c('prop.animal', 'prop.instrument', 'prop.other'), names_to="target_type", values_to="proportion")

# Summarize the condition-level data for a barplot.
first.move.orig.summary <- first.move.orig.subject.summary %>%
  group_by(condition, target_type) %>%
  summarize(M=mean(proportion), SE=sd(proportion)/sqrt(n())) %>% 
  mutate(bias = factor(condition, levels = c("Inst", "Equi", "mod") ))
```

```{r E4-mouse-moves-fig-web-and-orig, fig.cap = "Proportion of first mouse movements by location and verb bias in original dataset (Ryskin et al., 2017) ."}

ggplot(first.move.orig.summary, aes(x=bias, fill=target_type, y=M, ymin=M-SE,ymax=M+SE))+
  geom_col(position=position_dodge(width=0.9), color = "black")+
  geom_errorbar(position=position_dodge(width=0.9), width=0.2)+
  scale_fill_brewer(palette = "Set1", name = "Location", labels = c("Animal", "Instrument", "Other"))+
  coord_cartesian(ylim = c(0,1))+
  theme_classic()+
  labs(y = "Proportion of first mouse movements", x = NULL)
```