### 2022-04-01 group-d/E4-analysis-comparison-to-original.Rmd

---
title: "E4-analysis-comparison-to-original"
author: "Rachel Ryskin"
output: pdf_document
---

```{r , include=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
```

```{r E4-load-web-data}
first.move = read_csv("output/E4_mouse_data.csv")
eyetracking.window.summary.by.trial = read_csv( "output/E4_eye-tracking_data.csv") %>% 
  rename("condition" = compatibility) %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal) %>% 
  mutate(study = "web")
```

```{r load-data-from-original-study}

first.move.orig = read_tsv("original_study_data/Experiment1_clickData.txt")

eyetracking.window1.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_verb.txt") %>% 
  mutate(subject = as.character(subj), 
         prop.fixations.animal = TA/(TA+TI+CA+CI),
         #sum_dur = TA+TI+CA+CI,
         condition = factor(cond, levels = c(0, 2, 1, 9), labels = c("instrument", "equibiased", "modifier", "filler")),
         time.window = "post-verb-onset-pre-animal-onset",
         study = "original") %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal, study)

eyetracking.window2.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_animal.txt")%>% 
  mutate(subject = as.character(subj), 
         prop.fixations.animal = TA/(TA+TI+CA+CI),
         condition = factor(cond, levels = c(0, 2, 1, 9), labels = c("instrument", "equibiased", "modifier", "filler")),
         time.window = "post-animal-onset-pre-instrument-onset",
         study = "original") %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal, study)

eyetracking.window3.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_instrument.txt")%>% 
  mutate(subject = as.character(subj), 
         prop.fixations.animal = TA/(TA+TI+CA+CI),
         condition = factor(cond, levels = c(0, 2, 1, 9), labels = c("instrument", "equibiased", "modifier", "filler")),
         time.window = "post-instrument-onset",
         study = "original") %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal, study)

```

```{r E4-summarizing-first-mouse-moves-for-plotting-web}
# Summarize the data by subject, calculating proportion of trials the the first move was to the animal, instrument, or other.
first.move.web.subject.summary <- first.move %>%
  group_by(subject, compatibility) %>%
  summarize(prop.animal = mean(first.move.type == 'animal'),
            prop.instrument = mean(first.move.type == 'instrument'),
            prop.other = mean(first.move.type == 'other')) %>%
  pivot_longer(c('prop.animal', 'prop.instrument', 'prop.other'), names_to="target_type", values_to="proportion")

# Summarize the condition-level data for a barplot.
first.move.web.summary <- first.move.web.subject.summary %>%
  group_by(compatibility, target_type) %>%
  summarize(M=mean(proportion), SE=sd(proportion)/sqrt(n())) %>% 
  mutate(bias = factor(compatibility, levels = c("instrument", "equibiased", "modifier") ),
         study = "web")
```

```{r E4-summarizing-first-mouse-moves-for-plotting-orig}
# Summarize the data by subject, calculating proportion of trials the the first move was to the animal, instrument, or other.
first.move.orig.subject.summary <- first.move.orig %>%
  group_by(subject, condition) %>%
  summarize(prop.animal = mean(clickedTargetAnimalFirst),
            prop.instrument = mean(clickedTargetInstrumentFirst),
            prop.other = mean(clickedTargetAnimalFirst == 0 & clickedTargetInstrumentFirst == 0 )) %>%
  pivot_longer(c('prop.animal', 'prop.instrument', 'prop.other'), names_to="target_type", values_to="proportion")

# Summarize the condition-level data for a barplot.
first.move.orig.summary <- first.move.orig.subject.summary %>%
  group_by(condition, target_type) %>%
  summarize(M=mean(proportion), SE=sd(proportion)/sqrt(n())) %>% 
  mutate(bias = factor(condition, levels = c("Inst", "Equi", "mod"), labels = c("instrument", "equibiased", "modifier") ),
         study = "original")
```

The web version of the study qualitatively replicates the action and eye-tracking results of the original dataset [@ryskinVerbBiasesAre2017]. 
The mouse click results from both studies are summarized in Figure \@ref(fig:E4-mouse-moves-fig-web-and-orig).
The quantitative patterns of clicks were similar to those observed in the original dataset, though for Instrument-biased verbs, clicks were closer to evenly split between the animal and the instrument relative to the in-lab study where they were very clearly biased toward the instrument.

```{r E4-mouse-moves-fig-web-and-orig, fig.cap = "Proportion of first mouse movements by location and verb bias in the original dataset (Ryskin et al., 2017) and the current data collected online."}

ggplot(bind_rows(first.move.orig.summary, first.move.web.summary), 
       aes(x=bias, fill=target_type, y=M, ymin=M-SE,ymax=M+SE, alpha = study))+
  geom_col(position=position_dodge(width=0.9), color = "black")+
  geom_errorbar(position=position_dodge(width=0.9), width=0.2)+
  scale_fill_brewer(palette = "Set1", name = "Location", labels = c("Animal", "Instrument", "Other"))+
  scale_alpha_manual(values = c(0.5,1))+
  coord_cartesian(ylim = c(0,1))+
  theme_classic()+
  labs(y = "Proportion of first mouse movements", x = "Bias")
```

The eye-tracking results from both studies are summarized in Figure \@ref(fig:E4-proportion-fix-by-window-both). 
For simplicity, and to reflect the dependent variable used in analyses, we average the proportion of fixations to the target animal within each time window.
Though the qualitative patterns are replicated, proportions of fixations to the target animal were much lower in the web version of the study. 
This may reflect the fact that participants in the web study are less attentive and/or the quality of the webgazer eye-tracking system is lower, relative to the Eyelink 1000 which was used for the original study.

```{r}
eyetracking.window.summary.by.trial.both = bind_rows(eyetracking.window1.summary.by.trial.orig, eyetracking.window2.summary.by.trial.orig, eyetracking.window3.summary.by.trial.orig, eyetracking.window.summary.by.trial) %>% 
   filter(time.window != "end", time.window != "pre-verb-onset", condition != "filler" )
```

```{r}
eyetracking.window.summary.by.subj.both = bind_rows(eyetracking.window1.summary.by.trial.orig, eyetracking.window2.summary.by.trial.orig, eyetracking.window3.summary.by.trial.orig, eyetracking.window.summary.by.trial) %>% 
   filter(time.window != "end", time.window != "pre-verb-onset", condition != "filler" ) %>% 
  group_by(study, condition, time.window, subject) %>% 
  summarize(mean_prop = mean(prop.fixations.animal, na.rm = T)) %>% 
  mutate(time.window = factor(time.window, 
                            levels = c("post-verb-onset-pre-animal-onset",
                                       "post-animal-onset-pre-instrument-onset",
                                       "post-instrument-onset"),
                            labels = c('Verb to animal \n("Rub the")',
                                       'Animal onset\n("frog with the")',
                                       'Instrument+1500ms\n(feather...")')),
         condition  = factor(condition, 
                             levels = c("modifier", "equibiased", "instrument"), 
                             labels = c("Modifier", "Equi-biased", "Instrument"))
         ) %>% 
  ungroup()

eyetracking.window.summary.both = eyetracking.window.summary.by.subj.both %>% 
  group_by(study, condition, time.window) %>% 
  summarize(mean_prop = mean(mean_prop, na.rm = T), 
            se_prop = sd(mean_prop, na.rm = T)/sqrt(n()))

```

```{r E4-proportion-fix-by-window-both, fig.cap = "Proportion of target fixations by verb bias in the original dataset (Ryskin et al., 2017) and the current data collected online. Error bars reflect bootstrapped 95% CIs over subject means"}

ggplot(eyetracking.window.summary.by.subj.both)+
  stat_summary(aes(x=time.window, color=condition, y = mean_prop, shape = study), 
               geom = 'point', fun = 'mean', size = 3)+
 stat_summary(aes(x=as.numeric(time.window), y = mean_prop, linetype = study, group=paste(condition, study) ),
             geom = 'line', fun = 'mean', alpha = 0.5, color = "grey")+
  stat_summary(aes(x=as.numeric(time.window), y = mean_prop, color = condition, group = paste(condition , study)), 
                geom = 'errorbar', fun.data = 'mean_cl_boot', width = 0.05)+
  #facet_wrap(~time.window)+
  scale_color_brewer(palette = "Set1")+
  scale_shape_manual(values = c(1,16))+
  scale_linetype_manual(values = c("dashed","solid"))+
  coord_cartesian(ylim = c(0,.8))+
  theme_classic()+
  labs(y = "Proportion of fixations", x = NULL)
```
```{r, eval  = F}
orig.data.time.window.1 <- eyetracking.window1.summary.by.trial.orig %>% 
  filter(condition != "filler") %>% 
  mutate(condition = fct_drop(condition))
contrasts(orig.data.time.window.1$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))

model.time.window.1 <- lmer(prop.fixations.animal ~ condition + (1 | subject) + (1 | trialID), data=orig.data.time.window.1 )
summary(model.time.window.1)
```

```{r, eval  = F}
orig.data.time.window.2 <- eyetracking.window2.summary.by.trial.orig %>% 
  filter(condition != "filler") %>% 
  mutate(condition = fct_drop(condition))
contrasts(orig.data.time.window.2$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))

model.time.window.2 <- lmer(prop.fixations.animal ~ condition + (1 | subject) + (1 | trialID), data=orig.data.time.window.2 )
summary(model.time.window.2)
```

```{r, eval  = F}
orig.data.time.window.3 <- eyetracking.window3.summary.by.trial.orig %>% 
  filter(condition != "filler") %>% 
  mutate(condition = fct_drop(condition))
contrasts(orig.data.time.window.3$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))

model.time.window.3 <- lmer(prop.fixations.animal ~ condition + (1 | subject) + (1 | trialID), data=orig.data.time.window.3 )
summary(model.time.window.3)
```