### 2022-04-02 group-b/E2-analysis-calibration.Rmd
added E2 analyses to manuscript

---
title: "Group B Analysis - Calibration"
output: pdf_document
---

```{r message=FALSE}
library(jsonlite)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(lmerTest)
library(broom.mixed)
```

<!--Read JSON files and bind together into a tibble.-->

```{r}
data.files <- list.files('data', full.names = TRUE, pattern=".json")
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  return(data.table)
})
all.data <- bind_rows(data.tables)
```

```{r}

all.data.calib = all.data %>% 
  filter(trial_type == "webgazer-validate") %>% 
  dplyr::select(subject_id, trial_index,  percent_in_roi, average_offset) %>% 
  tidyr::unnest(percent_in_roi)

summary.data.calib = all.data.calib %>% 
  group_by(subject_id, trial_index) %>% 
  summarize(mean_percent_in_roi = mean(percent_in_roi)) %>% 
  group_by(subject_id) %>% 
  mutate(calib_num = row_number())
```

```{r, eval = F}
ggplot(summary.data.calib)+
  geom_line(aes(x = calib_num, y = mean_percent_in_roi, color=subject_id))+
  theme_bw()+
  theme(legend.position = "none") 
```
```{r, eval=F}
summary.data.calib.wide = summary.data.calib %>% 
  select(-trial_index) %>% 
  pivot_wider(id_cols = subject_id, names_from=calib_num, values_from = mean_percent_in_roi )

# correlation between initial and halfway calibration
ggplot(summary.data.calib.wide %>% filter(is.na(`3`)), aes(x = `1`, y = `2`))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()+
  theme(legend.position = "none")

# correlation between 2 successive calibration attempts
ggplot(summary.data.calib.wide %>% filter(!is.na(`3`)), aes(x = `1`, y = `2`))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()+
  theme(legend.position = "none")

# correlation between second attempt calibration and halfway
ggplot(summary.data.calib.wide %>% filter(!is.na(`3`)), aes(x = `2`, y = `3`))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()+
  theme(legend.position = "none") 
```

```{r}

calib.by.subj = summary.data.calib %>% 
  group_by(subject_id) %>% 
  summarize(mean_percent_in_roi = mean(mean_percent_in_roi))

eyetracking.effects.by.subj = read_csv( "output/E2_eye-tracking_data_subj.csv") %>% 
 # rename("condition" = compatibility) %>% 
  filter(condition == "fixed" ) %>% 
  group_by(subject_id, normalized_quadrant) %>% 
  summarize(M = mean(proportion)) %>% 
  pivot_wider(names_from = normalized_quadrant, values_from = M) %>% 
  mutate(crit_bias = critical - (first+second+third)/3 ) %>% 
  left_join(calib.by.subj, by = "subject_id")

```

```{r, eval = F}
ggplot(eyetracking.effects.by.subj, aes(x = mean_percent_in_roi, y = crit_bias))+
  geom_point()+
  geom_smooth(method = "lm")
```
```{r}

m_calib = broom::tidy(cor.test(eyetracking.effects.by.subj$mean_percent_in_roi, eyetracking.effects.by.subj$crit_bias))
#m_calib 
```

```{r}
m_calib_r = m_calib %>% pull(estimate ) 
m_calib_p = m_calib %>% pull(p.value )

```

Participants' calibration quality, measured as the mean percentage of fixations that landed within `r colorize("XX", "red")` pixels of the calibration point, varied substantially (between `r range(calib.by.subj$mean_percent_in_roi) |> round(digits = 2)` %). 
The quality of a participant's calibration was not significantly correlated with the participant's effect size ( _Pearson's r_= `r m_calib_r`, _p_ = `r m_calib_p`) as measured by the difference between the proportion of looks to the critical quadrant minues the average proportion of looks to the average of the other three quadrants.