### 2022-04-04 group-c/E3-analysis-roi.Rmd
add E3 ROI analysis

---
title: "Experiment 3: Replication Analysis"
output: html_notebook
---

In the original experiment, the two objects on day 1 were presented on two separate monitors and gaze was coded by manually coding video recordings. In our replication analysis, we analyzed eye movement data using ROIs. In this section we explore an alternative coding of the eye movement data by coding simply left half vs. right half of the screen.

```{r E3 load libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(jsonlite)
library(ggplot2)
```

```{r E3 load all data, include=FALSE}
data.files <- list.files('data', full.names = TRUE)
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  data.table$response <- as.character(data.table$response)
  return(data.table)
})
all.data <- bind_rows(data.tables)
```

```{r Split into day 1 and day 2, include=FALSE}
study.id.day.1.group.1 <- "6062337f1a63a6788185d0e3"
study.id.day.1.group.2 <- "607de170bb44b7e25174421f"
study.id.day.2.group.1 <- "606233c0e42b980319b6a18e"
study.id.day.2.group.2 <- "607de2837874a824e2e8c6ec"

all.data <- all.data %>%
  mutate(day = if_else(study_id %in% c(study.id.day.1.group.1, study.id.day.1.group.2), 1, 2)) %>%
  mutate(group = if_else(study_id %in% c(study.id.day.1.group.1, study.id.day.2.group.1), 1, 2))

data.day.1 <- all.data %>% filter(day == 1)
data.day.2 <- all.data %>% filter(day == 2)
```

```{r Select eyetracking data, include=FALSE}
eyetracking.data <- data.day.1 %>%
  filter(phase == "part2") %>%
  select(subject_id, webgazer_data, webgazer_targets, left_image_new, right_image_new, right_image, left_image) %>%
  unpack(webgazer_targets) %>%
  unpack(c("#left-img", "#right-img"), names_sep=".") %>%
  unnest(webgazer_data)
```

```{r Add ROI information, include=FALSE}
in.box <- function(x, y, left, right, top, bottom, padding){
  is.in.the.box <- x >= left - padding & x <= right + padding & y >= top - padding & y <= bottom + padding
  return(is.in.the.box)
}

eyetracking.data <- eyetracking.data %>%
  mutate(midpoint.x = `#left-img.right` + (`#right-img.left` - `#left-img.right`)/2) %>%
  mutate(in.left.roi = x < midpoint.x) %>%
  mutate(in.right.roi = x > midpoint.x)%>%
  mutate(in.target.roi = if_else(left_image_new == TRUE, in.left.roi, in.right.roi))%>%
  mutate(in.distractor.roi = if_else(left_image_new == TRUE, in.right.roi, in.left.roi))
```

```{r Calculate looking score, include=FALSE}
looking.score.data <- eyetracking.data %>% 
  group_by(subject_id, right_image, left_image, right_image_new, left_image_new) %>%
  summarise(looking.score = if_else(sum(in.target.roi, in.distractor.roi) == 0, 0, sum(in.target.roi)/ sum(in.target.roi, in.distractor.roi))) %>%
  mutate(image = if_else(right_image_new, left_image, right_image))
```

```{r Day 1 t-test looking scores, include=FALSE}
t.test.data1 <- looking.score.data %>%
  group_by(subject_id) %>%
  summarize(M.looking.score = mean(looking.score))

t.test.day1 <- t.test(t.test.data1$M.looking.score, mu = 0.5)
```

#### Looking Scores

When looking scores are coded as left vs. right half of the screen, we do find that participants looked more at the novel object. The mean looking score was `r mean(t.test.data1$M.looking.score)` (*SD* = `r sd(t.test.data1$M.looking.score)`). This was significantly greater than 0.5, *t*(`r t.test.day1$parameter`) = `r t.test.day1$statistic`, *p* = `r t.test.day1$p.value`.

#### Correlations with Day 2 Performance

```{r Prepare Day 2 data, include=FALSE}
day2.test.data <- data.day.2 %>%
  filter(task %in% c("recognition", "confidence")) %>%
  select(subject_id, image, old_or_new, correct, rt, response, task) %>%
  pivot_wider(names_from=task, values_from=c(correct, response, rt)) %>%
  select(-correct_confidence) %>%
  mutate(response_confidence = as.numeric(response_confidence))
```

```{r Calculate Looking Score Correlations, include=FALSE}
looking.plus.behavioral <- left_join(looking.score.data, day2.test.data)

looking.confidence.correlations.subjects <- looking.plus.behavioral %>%
  filter(correct_recognition == TRUE) %>%
  filter(rt_recognition <= 10000) %>%
  group_by(subject_id) %>%
  summarize(pearson.r = cor(looking.score, response_confidence)) %>%
  mutate(r.to.z = .5*(log(1+pearson.r) - log(1 - pearson.r)))

looking.confidence.correlations.subjects.filtered <- looking.confidence.correlations.subjects %>%
  filter(!is.na(pearson.r))

looking.confidence.correlation.t.test <- t.test(looking.confidence.correlations.subjects.filtered$r.to.z)

looking.rt.correlations.subjects <- looking.plus.behavioral %>%
  filter(correct_recognition == TRUE) %>%
  filter(rt_recognition <= 10000) %>%
  group_by(subject_id) %>%
  summarize(pearson.r = cor(looking.score, rt_recognition)) %>%
  mutate(r.to.z = .5*(log(1+pearson.r) - log(1 - pearson.r)))

looking.rt.correlations.subjects.filtered <- looking.rt.correlations.subjects %>%
  filter(!is.na(pearson.r))

looking.rt.correlation.t.test <-t.test(looking.rt.correlations.subjects.filtered$r.to.z)
```

Performance on day 2 remained uncorrelated with day 1 looking scores after switching the coding of gaze. We found no significant different from 0 for the correlation between looking score and confidence ratings, t(`r looking.confidence.correlation.t.test$parameter`) = `r looking.confidence.correlation.t.test$statistic`, p = `r looking.confidence.correlation.t.test$p.value` (excluding the subjects who gave the same confidence judgment for all images), nor the the correlation between looking score and RT, t(`r  looking.rt.correlation.t.test$parameter`) = `r looking.rt.correlation.t.test$statistic`, p = `r looking.rt.correlation.t.test$p.value`.