### 2022-08-24 group-c/E3-calibration-figure.Rmd
some work on group C / exp 3

---
title: "Experiment 3: Calibration Figure"
output: html_notebook
---

```{r include=FALSE}
library(ggplot2)
library(dplyr)
```

```{r include=FALSE}
calibration_points <- matrix(c(
  c(10,50),
  c(50,50),
  c(90,50),
  c(25,25),
  c(25,75),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)

colnames(calibration_points) <- c("x","y")

df <- as_tibble(calibration_points)
```

```{r echo=FALSE, fig.cap="Calibration point locations for Experiment 3."}
ggplot(df, aes(x=x, y=y))+
  geom_point(size=3)+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  theme(panel.border = element_rect(color="black", fill=NA))
```