### 2022-09-25 group-b/E2-calibration-figure.Rmd

---
title: "Experiment 2: Calibration Figure"
output: html_notebook
---

```{r include=FALSE}
library(ggplot2)
library(dplyr)
```

```{r include=FALSE}
# calib: [25,25],[25,75],[50,50],[75,25],[75,75]
calibration_points <- matrix(c(
  c(25,25),
  c(25,75),
  c(50,50),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)
# valid: c(25, 25, 50, 75, 75, 25, 75, 50, 25, 75)
#      [,1] [,2]
# [1,]   25   25
# [2,]   25   75
# [3,]   50   50
# [4,]   75   25
# [5,]   75   75
validation_points <- matrix(c(
  c(25,25),
  c(25,75),
  c(50,50),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)

colnames(calibration_points) <- c("x","y")
colnames(validation_points) <- c("x","y")

df <- as_tibble(calibration_points) %>% bind_rows(as_tibble(validation_points))
df$point <- c(rep("calibration", nrow(calibration_points)), rep("validation", nrow(validation_points)))
```

```{r echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 2. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.(In this experiment all the same locations were used for both calibration and validation.)", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```