### 2022-09-25 manuscript.Rmd

---
title             : "What paradigms can webcam eye-tracking be used for? Attempted replications of 5 “classic” cognitive science experiments"
shorttitle        : "Webcam eye-tracking paradigms"

author:
  - name: Joshua R. de Leeuw
    affiliation: '1'
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Investigation
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    corresponding: yes
    email: jdeleeuw@vassar.edu
    address: 124 Raymond Ave, Poughkeepsie, NY 12604, USA
  - name: Rachel Ryskin
    affiliation: '2'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: rryskin@ucmerced.edu
  - name: Ariel N. James
    affiliation: '3'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: ajames2@macalester.edu
  - name: Joshua K. Hartshorne
    affiliation: '4'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: joshua.hartshorne@bc.edu
  - name: Haylee Backs
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: hbacks@vassar.edu
  - name: Nandeeta Bala
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: nbala@vassar.edu
  - name: Laila Barcenas-Meade
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: lbarcenasmeade@vassar.edu
  - name: Samata Bhattarai
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: sbhattarai@vassar.edu
  - name: Tessa Charles
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: tcharles@vassar.edu
  - name: Gerasimos Copoulos
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: gcopoulos@vassar.edu
  - name: Claire Coss
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: ccoss@vassar.edu
  - name: Alexander Eisert
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: aeisert@vassar.edu
  - name: Elena Furuhashi
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: efuruhashi@vassar.edu
  - name: Keara Ginell
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: kginell@vassar.edu
  - name: Anna Guttman-McCabe
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: aguttmanmccabe@vassar.edu
  - name: Emma (Chaz) Harrison
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: eharrison@vassar.edu
  - name: Laura Hoban
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: lhoban@vassar.edu
  - name: William A. Hwang
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: whwang@vassar.edu
  - name: Claire Iannetta
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: ciannetta@vassar.edu
  - name: Kristen M. Koenig
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: kkoenig@vassar.edu
  - name: Chauncey Lo
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: chaunceylo@vassar.edu
  - name: Victoria Palone
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: vpalone@vassar.edu
  - name: Gina Pepitone
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: gpepitone@vassar.edu
  - name: Margaret Ritzau
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: mritzau@vassar.edu
  - name: Yi Hua Sung
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: ysung@vassar.edu
  - name: Lauren Thompson
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: lthompson@vassar.edu

affiliation:
  - id: '1'
    institution: Cognitive Science Department, Vassar College
  - id: '2'
    institution: Department of Cognitive & Information Science, University of California,
      Merced
  - id: '3'
    institution: Psychology Department, Macalester College
  - id: '4'
    institution: Department of Psychology & Neuroscience, Boston College

abstract: |
  ADD LATER

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "eye-tracking, online, webcam, jsPsych, cognitive science"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
r_refs("references.bib")

colorize <- function(x, color) { # for adding colored text (e.g., for notes to self)
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{=html}
<!-- MAJOR TO DOs: 
 - get data from in-lab versions of group A (AJ) and group D (RR)
 - replicate group D and A in lab

 (see bottom for notes from e-mails)
 -->
```
Intro stuff:

-   Eye-tracking as a key method in cognitive science research

-   Online data collection is more and more popular & let's us ask new
    questions, test more diverse populations

-   But, concerns over quality + little known about eye-tracking online

# Present work

In order to validate online eyetracking measures, we set out to reproduce five previously published studies representing a variety of questions, topics, and paradigms. The goal was to examine the strengths and weaknesses of webcam eye-tracking for common paradigms in cognitive science. Ideally, we would only attempt to replicate studies where the original measurements have small error bars and are known to replicate; otherwise, it can be difficult to distinguish a failure of the method (online eyetracking does not work) from a failure of the original study to replicate.

In practice, replications (successful or otherwise) have only been reported for a small number of studies, so we ultimately included some studies with unknown replicability. We addressed this in several ways. First, replicating five very different studies from different research traditions decreases our reliance on any one study. Second, we include several ``sanity check'' analyses, such as the correlation between calibration accuracy and effect size. (If the effect is real but there is noise from low-accuracy eyetracking, this correlation should be substantial.) Third, for XXX of the studies, we had comparison data collected in-lab either using jsPsych or a more traditional eyetracker technology, allowing us to direct assess the impact of differences in subject population and equipment.

## Selection of Studies

<!-- June 2022: shimojo: 964 citations, altmann: 1840, ryskin: 75, johansson 190, 127 -->
<!-- RR: switched ryskin to Snedeker & trueswell 2004 bcs my 75 looked sad -->

We chose five high-impact eyetracking studies involving adult subjects. (Given the additional difficulties of recruiting and retaining child participants, we excluded developmental studies.) Our goal was to include experiments from a range of topic areas (e.g., memory, decision making, psycholinguistics) and paradigms (two
halves of the screen, visual world paradigm with four quadrants, visual
world paradigm with "natural" scenes). As noted above, we had a preference for well-established findings that are known to replicate, though for sake of diversity this was not always possible. Table\ \ref(tab:studies-table) provides an overview of the five studies we selected.

```{r studies-table, echo=FALSE}
studies_table <- data.frame(
  Citation = c(
    "Altmann & Kamide, 1999",
    "Johansson & Johansson, 2013",
    "Manns, Stark, & Squire, 2000",
    "Snedeker & Trueswell, 2004",
    "Shimojo et al., 2003"
  ),
  `Topic Area` = c(
    "Psycholinguistics",
    "Memory",
    "Memory",
    "Psycholinguistics",
    "Decision Making"
  ),
  `Paradigm` = c(
    "Natural Scenes",
    "Four Quadrants",
    "Two Halves",
    "Four Quadrants",
    "Two Halves"
  ),
  `Citations (June 2022, Google Scholar)` = c(
    1840,
    190,
    127,
    448,
    964
  ),
  check.names = FALSE
)

apa_table(
  studies_table,
  caption = "Studies selected for replication attempts",
  font_size = "small"
)

```

# General Methods

## Participants

Participants completed the experiment remotely and were recruited
through the Prolific platform. In order to have access to the
experiment, participants had to meet the following criteria: 18 years of
age or older, fluency in English, and access to a webcam. All
participants provided informed consent. The studies were approved by the
Vassar College Institutional Review Board.

In order to have adequate statistical power and precision, we aimed for
2.5x the sample size of the original experiment, following the heuristic
of Simonsohn [@simonsohn2015]. In study 5, the original sample size was so
small that we opted to collect 5x the number of participants to increase
precision. Because of budget and time constraints we were unable to
replace the data for subjects who were excluded or whose data was
missing due to technical failures.

## Equipment

We used a fork of the `webgazer.js` library for webcam eyetracking
[@papoutsaki2016webgazer], implemented in `jsPsych`, a Javascript
library for running behavioral experiments in a web browser
[@deleeuwJsPsychJavaScriptLibrary2015]. Our fork included changes to
webgazer.js in order to improve data quality for experiments in which
the precise timing of stimulus onsets is relevant. Specifically, we
implemented a polling mode so that gaze predictions could be requested
at a regular interval, which improved the sampling rate considerably in
informal testing. This modification is similar to what Yang and Krajbich
[-@yang2021webcam] reported improved the sampling rate in their study of
webgazer. We also adjusted the mechanism for recording time stamps of
each gaze prediction, so that the time stamp reported by webgazer is
based on when the video frame is received and not when the computation
of the gaze point is finished.

## Eye-tracking Calibration and Validation

<!--took paragraph below from group D's methods -->

When participants began the experiment, they were notified the webcam
would be used for eye tracking but no video would be saved. They were
asked to remove glasses if possible, close any other tabs or apps, turn
off notifications, and make sure their face was lit from the front. The
webcam's view of the participant popped up on the screen, and
participants were asked to center their face in the box and keep their
head still. The experiment window then expanded to full screen, and
participants began the eye-tracking calibration.

During the calibration, dots appeared on the screen one at a time in
different locations, and the participants had to fixate them and click
on each one. Once they clicked on a dot, it would disappear and a new
one would appear in a different location on the screen. The locations of
calibration dots were specific to each experiment (details below) and
appeared in the areas of the screen where the visual stimuli would
appear during the main task in order to ensure that eye movements were
accurately recorded in the relevant regions of interest. After the
calibration was completed, the validation began. Participants were asked
to go through the same steps as the calibration, except that they only
fixated the dots as they appeared in different locations on the screen.
If accuracy on the validation was too low (fewer than 50% of looks
landed within a 200 px radius of the validation points<!--check!-->),
participants were given an opportunity to re-start the calibration and
validation steps. If the second attempt also lead to low validation
accuracy, participants were informed that they could not participate in
the study.

## Data pre-processing

We used `r cite_r("references.bib")` for all our analyses.

# Experiment 1

<!-- group a -->

The first study was a replication attempt of
@altmannIncrementalInterpretationVerbs1999. Altmann and Kamide used the
visual world eye-tracking paradigm [@tanenhaus1995] to show that
meanings of verbs rapidly constrain the set of potential subsequent
referents in sentence processing. For example, when looking at the
display in Figure\ \@ref(fig:E1-example-trial) and listening to a sentence like
"The boy will eat the...," participants are more likely to look at the cake than 
when they hear "The boy will move the...," in which case they tend to look at the
train, presumably because cakes are edible and trains are not. Semantic
information available at the verb is used to anticipate upcoming
linguistic input.

```{r read_groupA_data}

```

## Methods

<!-- since there are not the analysis scripts we're using, I'm not sure about the best way to link to this -->

All stimuli, experiment scripts, data, analysis scripts, and a
pre-registration are available on the Open Science Framework at
<https://osf.io/s82kz>.

### Participants

60 participants were paid \$2.60 for their participation. Our sample size
of participants was determined by the total run time of our experiment,
\~10 minutes, and the allotted funding from the Vassar College Cognitive
Science Department. From this information, we calculated a reasonable
number of participants we could afford to compensate on Prolific. Note
that the sample size of the original study was 24. For unknown reasons,
2 of the subjects' results were not recorded, so in the analysis, we
worked with data collected from 58 participants.

### Procedure

The task began with a 9-point eye-tracker calibration
and validation (Figure\ \@ref(fig:E1-calibration-figure)). 
During the experiment, the participants were
simultaneously presented with a visual image and a corresponding audio
recording of a spoken sentence. Participants had to input a keyboard
response indicating "yes" or "no" as to whether the sentence they heard
was feasible given the visual image. There were two practice trials to
ensure that participants understood the instructions
before they undertook the main portion of the experiment. Participants'
reaction times, keyboard responses, and looks to objects in the scene
were recorded for each trial.

```{r E1-calibration-figure, child='group-a/E1-calibration-figure.Rmd'}

```

```{=html}
<!-- Qs: 
-keyboard keys? what was the exact phrasing of the instruction? 
-Any delay between when visual scene appears and the audio starts? 
-Does trial end as soon as participant responds?-->
```
### Materials & Design

The visual stimuli were created through Canva and depicted an agent
accompanied by four to five objects in the scene (see Figure\ \@ref(fig:E1-example-trial)). On
critical trials, participants heard one of two sentences associated with
the scene. In the restrictive condition, the sentence (e.g., "The boy
will eat the cake") contained a verb (e.g., "eat") which restricts the
set of possible subsequent referents (e.g., to edible things). Only the
target object (e.g., the cake) was semantically consistent with the
verb's meaning. In the non-restrictive condition, the sentence (e.g.,
"The boy will move the cake") contained a verb (e.g., "move") which does
not restrict the set of possible subsequent referents. The target object
(e.g., the cake) as well as the distractor objects (e.g., the train, the
ball, etc.) were semantically consistent with the verb's meaning. Both
sentences were compatible with the scene, such that the correct keyboard
response for the critical trials was "yes." Filler trials consisted of
scenes that looked similar to critical scenes but were paired with
inappropriate sentences. The correct keyboard response for the filler
trials was "no."

```{r E1-example-trial, fig.cap='Example trial from Experiment 1. Participants would hear a sentence (e.g., "The boy will eat the cake") and respond according to whether the sentence matched the picture.'}
knitr::include_graphics("group-a/E1-example-trial.jpeg")
```

Each participant was presented with sixteen critical trials (eight in
the restrictive condition, eight in the non-restrictive condition) and
sixteen fillers for a total of 32 trials. The order of trials and the
assignment of critical scene to condition was random on a
subject-by-subject basis.<!-- right?-->

### Data pre-processing and analysis

Looks to the objects in the scene were time-locked to the onset of the
verb, the offset of the verb, onset of the post-verbal determiner, and
onset of the target noun.

## Results

### Replication

-   here we will describe the analyses that are as close as possible to
    the original paper with a minimal validation cutoff

-   same analysis but with stricter validation cutoff

### Comparison to in-lab data

-   here we will describe a direct comparison to data collected in the
    lab

### Calibration

-   here we will describe the analyses that correlate calibration
    quality with effect size at the individual level

## Discussion

# Experiment 2

<!-- group b -->

The second study was a replication attempt of @johanssonLookHereEye2014,
which examined how visuospatial information is integrated into memory
for objects. They found that, during memory retrieval, learners
spontaneously look to blank screen locations where pictures were located
during encoding [see @spiveyOculomotorMechanismsActivated2001] and that
this spatial reinstatement facilitates retrieval of the picture.

## Methods

All stimuli, experiment scripts, data, analysis scripts, and a
pre-registration are available on the Open Science Framework at
<https://osf.io/xezfu/>.

### Participants

60 participants were paid for their participation.
The sample size was motivated in part by budget constraints, but was
nonetheless 2.5x larger than the original sample size of 24). 
Data from 1 participant were not properly
recorded due to unknown technical issues, so data from 59 participants
were included in all analyses to follow.

### Procedure

The task began with a 9-point eye-tracker calibration
and validation (Figure\ \@ref(fig:E2-calibration-figure)).

```{r E2-calibration-figure, child='group-b/E2-calibration-figure.Rmd'}

```

The experiment consisted of two blocks each composed of an encoding
phase and a recall phase. During the encoding phase, participants saw a
grid indicating the four quadrants of the screen. Each quadrant
contained six images of items belonging to the same category (see Figure\ \@ref(fig:E2-example-trial)). 
The four categories were humanoids, household objects, animals, and
methods of transportation. 
Each of the four quadrants was presented one at a time. First, a list of
the items in the quadrant was shown, then the pictures of items were displayed in the quadrant. 
For each item, participants used their arrow keys to indicate whether the object was facing left or right. After the participant identified the direction of each
item, they would have an additional 30 seconds to encode the name and
orientation of each item in the quadrant. Finally, after all four quadrants
were presented, participants were shown the full grid of
24 items and had 60 seconds to further encode the name and orientation
of each item.

```{r E2-example-trial, fig.cap='Example trial from Experiment 2.'}
knitr::include_graphics("group-b/E2-example-figure.jpeg")
```

During the recall phase, participants listened to statements and
responded by pressing the 'F' key for false statements and 'T' for true
ones. Each statement fell into either an interobject or intraobject
condition. Interobject statements were those that compared two different
items in the grid (e.g. "The skeleton is to the left of the robot"),
while intraobject statements were those that asked about the orientation
of a single item (e.g. "The bus is facing right"). There were 48 total
statements, with 24 interobject and 24 intraobject statements split
evenly among the four quadrants. While listening to these statements, in
the free-viewing block, participants saw a blank screen and were allowed
to freely gaze around the screen. During the fixed-viewing block,
participants were asked to fixate a small cross in the center of the
screen throughout the recall phase. In both cases, the mouse was
obscured from the screen. Participants were randomly assigned to see the
fixed-viewing or free-viewing block first. Different images were used in each block.

After completing both encoding-recall blocks, participants were asked to
answer a few survey questions (such as whether they wore glasses or
encountered any distractions).

The primary methodological difference between this replication and
Johansson and Johansson's study was that the original study included two
additional viewing conditions that were omitted from this replication
due to time constraints. In those two conditions, participant were
prompted to look to a specific quadrant (rather than free viewing or
central fixation) which either matched or mismatched the original
location of the to-be-remembered item.

## Results

```{r E2-analysis-replication, child='group-b/E2-analysis-replication.Rmd'}

```

### Calibration

```{r E2-analysis-calibration, child='group-b/E2-analysis-calibration.Rmd'}

```

## Discussion

As in @johanssonLookHereEye2014 and @spiveyOculomotorMechanismsActivated2001, during memory retrieval, learners spontaneously look to blank screen locations where pictures were located
during encoding, suggesting that visuospatial information is integrated into the memory
for objects. However, we did not observe a memory benefit, in terms of speed or accuracy, of spatial reinstatement via gaze position during retrieval of the picture. We can speculate that this may be due to the fact that participants struggled to maintain their gaze fixed in the center in the fixed-viewing condition, such that the difference between the fixed- and free-viewing conditions was minimal. Crucially for the current purposes, the webcam-based eye-tracking measurements were successful in replicating the key eye-tracking results.

# Experiment 3

<!-- group c -->

The third study was a partial replication attempt of
@mannsVisualPairedcomparisonTask2000. This experiment used the visual
paired-comparison, which involves presenting a previously-viewed image
and novel image together and measuring the proportion of time spent
looking at each image. The expected pattern of results is that
participants will look more at novel objects. They
@mannsVisualPairedcomparisonTask2000 hypothesized that this pattern of
behavior could be used to measure the strength of memories. If a viewer
has a weak memory of the old image, then they may look at the old and
new images roughly the same amount of time. They tested this in two
ways. First, they showed participants a set of images, waited five
minutes, and then paired those images with novel images. They found that
participants spent more time (58.8% of total time) looking at the novel
images. They then measured memory performance one day later and found
that participants were more likely to recall images that they had spent
less time looking at during the visual paired-comparison task the
previous day.

## Methods

The stimuli, experimental code, and data and analysis scripts can be
found on the Open Science Framework at <https://osf.io/k63b9/>.
The pre-registration for the study can be
found at <https://osf.io/48jsv> . We
inadvertently did not create a formal pre-registration using the OSF
registries tool, but this document contains the same information and is
time stamped prior to the start of data collection. <!-- Students failed
to pre-register this one and I missed it. There's an unofficial
pre-registration in the OSF site. -->

### Participants

Our pre-registered target was 50 participants. 51 participants completed
the first day of the experiment and 48 completed the second day.
Following Manns et al., we excluded 3 participants due to perfect
performance on the recognition memory test because this prevents
comparison of gaze data for recalled vs. non-recalled images. Our final
sample size was 45 participants.

### Procedure

The task began with a 7-point eye-tracker calibration (each point was
presented 3 times in a random order) and validation with 3 points (each
presented once). The point locations were designed to focus calibration
on the center of the screen and the middle of the left and right halves
of the screen (Figure\ \@ref(fig:E3-calibration-figure)).

```{r E3-calibration-figure, child='group-c/E3-calibration-figure.Rmd'}

```

The experiment was administered over the course of two
consecutive days. It consisted of three sections: a presentation phase,
a test phase, and a recognition test. The first two phases occurred on
the first day, while the recognition test occurred on the second day.

During the presentation phase, participants viewed 24 pairs of identical
color photographs depicting common objects. Each pair was presented for
5 seconds and an interval of 5 seconds elapsed before the next pair was
shown. The order of the photographs was randomized and different for
each participant. After completion of the presentation phase,
participants were given a 5-minute break during which they could look
away from the screen.

After the break, they were prompted to complete the eye-tracking
calibration again before beginning the test phase. During this phase,
participants again viewed 24 pairs of photographs with an interstimulus
duration of 5 seconds. In each pair, one photograph was previously seen
during the presentation phase, while the other was new. Which pictures
were old or new was counterbalanced across participants.<!-- 2 lists?-->
For half of the participants in each counterbalancing group, the new and
old photographs were
reversed.<!-- not sure what that last sentence means: is it about the side of the screen? or still just describing that there were 2 lists-->

Approximately 24 hours after completing the first session, with a leeway
interval of 12 hours to accommodate busy schedules, participants were
given the recognition test. It consisted of 48 photographs, presented
one at a time. Each was shown on the screen for 1 second, followed by a
1 second interstimulus interval. Half of the photographs had been
viewed twice on the previous day and were deemed the "targets." The
other half depicted an object with the same name as an object in one of
the old photographs, but had not been viewed before, deemed "foils."
Each photograph remained on the screen until the participants indicated
whether or not they had seen it before by pressing 'y' for yes and 'n'
for no. After they pressed one of the two keys, a prompt on the screen
asked them to rate their confidence in their answer from 1 as a "pure
guess" to 5 as "very sure." by clicking on the corresponding number on
the screen. No feedback on their responses was given during the test.

The experimental design is visually depicted in Figure\ \@ref(fig:E3-design-schematic)

```{r E3-design-schematic, fig.cap='Schematic of the design of Experiment 3'}
knitr::include_graphics("group-c/E3-example-figure.jpeg")
```

### Materials

Images were selected XXX...

There were two modifications we made to the methods of the original
experiment. As we are only replicating the declarative memory component
of the original experiment, we did not have a "priming group."
Therefore, we followed only the procedure for the "looking group."
Additionally, for each section of the study, the stimuli was presented
on a single screen instead of two screens due to the constraints of the
online experiment format.

## Results

```{r E3-analysis-replication, child='group-c/E3-analysis-replication.Rmd'}

```

### Effects of ROIs

```{r E3-analysis-ROIs, child='group-c/E3-analysis-roi.Rmd'}

```

### Calibration

```{r E3-analysis-calibration, child='group-c/E3-analysis-calibration.Rmd'}

```

## Discussion

# Experiment 4

<!-- group d -->

The fourth study was a replication attempt of Experiment 1 in
@ryskinVerbBiasesAre2017, which was closely modeled on
@snedekerDevelopingConstraintsParsing2004a. These studies used the
visual world paradigm to show that listeners use knowledge of the
co-occurrence statistics of verbs and syntactic structures to resolve
ambiguity. For example, in a sentence like "Feel the frog with the
feather," the phrase "with the feather" could be describing the frog, or
it could be describing the instrument that should be used to do the
"feeling." When both options (a frog holding a feather and a feather by
itself) are available in the visual display, listeners rely on the
verb's "bias" (statistical co-occurrence either in norming or corpora)
to rapidly choose an action while the sentence is unfolding. .

## Methods

The stimuli, experimental code, and data and analysis scripts can be
found on the Open Science Framework at the following link,
<https://osf.io/x3c49/>. The pre-registration
for the study can be found at <https://osf.io/3v4pg>.

### Participants

57 participants were paid \$2.50 for their participation. A sample size of 60 was initially chosen (but not reached in time) because we wanted to replicate the experiment with greater statistical power. Note that the original study had a sample size of 24.

### Procedure

After the eye-tracking calibration and validation (Figure\ \@ref(fig:E4-calibration-figure)), participants went through an audio
test so they could adjust the audio on their computer to a comfortable
level. Before beginning the experiment, they were given instructions
that four objects would appear, an audio prompt would play, and they
should do their best to use their mouse to act out the instructions.
They then went through three practice trials which were followed by 54
critical trials and 24 filler trials presented in a random order.

```{r E4-calibration-figure, child='group-d/E4-calibration-figure.Rmd'}

```

During a trial, four pictures were displayed (target animal, target
instrument, distractor animal, distractor instrument), one in each
corner of the screen, and participants heard an audio prompt that
contained instructions about the action they needed to act out (e.g.,
"Rub the butterfly with the crayon"; see Figure\ \@ref(fig:E4-example-trial))[^1]. Using their
cursor, participants could act out the instructions by clicking on
objects and moving them or motioning over the objects[^2]. After the
action was completed, the participants were instructed to press the
space bar which led to a screen that said "Click Here" in the middle in
order to remove bias in the eye and mouse movements from the previous
trial. The experiment only allowed the participants to move on to the
next trial once the audio was completely done playing and the mouse had
been moved over at least one object.

[^1]: In the original study, the pictures appeared one by one on the
    screen and their names were played as they appeared. We removed this
    introductory portion of the trial to save time

[^2]:  As opposed to the original study we recorded mouse movement
    instead of clicking behavior since not all of the audio prompts
    required clicking. For example, the sentence "locate the camel with
    the straw" may not involve any clicking but rather only mousing over
    the camel.

```{r E4-example-trial, fig.cap='An example of a critical trial from Experiment 4 for the sentence “Rub the butterfly with the crayon.” The butterfly is the target animal, the panda is the distractor animal, the crayon is the target instrument, and the violin is the distractor instrument.'}
knitr::include_graphics("group-d/E4-example-figure.jpeg")
```

### Materials

The images and audios presented to the participants were the same
stimuli used in the original study (available here <link>). The critical
trials were divided into modifier-biased, instrument-biased, and
equibiased conditions, and the filler trials did not contain ambiguous
instructions. Two lists of critical trials were made with different verb
and instrument combinations (e.g., "rub" could be paired with "panda"
and "crayon" in one list and "panda" and "violin" in the second list).
Within each list, the same verb was presented twice but each time with a
different target instrument and animal. The lists were randomly assigned
to the participants to make sure the effects were not caused by the
properties of the animal or instrument images used. The list of verbs
used can be found in Appendix A of the original study.

## Results

```{r E4-analysis-replication, child='group-d/E4-analysis-replication.Rmd'}

```

### Comparison to in-lab data

```{r E4-analysis-comparison-to-original-study, child='group-d/E4-analysis-comparison-to-original.Rmd'}

```

### Calibration

```{r E4-analysis-calibration, child='group-d/E4-analysis-calibration.Rmd'}

```

### Effects of ROIs

```{r E4-analysis-ROIs, child='group-d/E4-analysis-ROIs.Rmd'}

```

## Discussion

# Experiment 5

<!-- group e -->

The fifth study was a replication attempt of @shimojoGazeBiasBoth2003,
which found that human gaze is actively involved in preference
formation. Separate sets of participants were shown pairs of human faces
and asked either to choose which one they found more attractive or which
they felt was rounder. Prior to making their explicit selection,
participants were increasingly likely to be fixating the face they
ultimately chose, though this effect was significantly weaker for
roundness discrimination.

Note that Shimojo and colleagues compare five conditions, of which we
replicate only the two that figure most prominently in their
conclusions: the "face-attractiveness-difficult task" and the
"face-roundness task".

## Methods

All stimuli, experiment scripts, data, and analysis scripts are
available on the Open Science Framework at <https://osf.io/eubsc/>
(<https://osf.io/eubsc/>). The study pre-registration is available at
<https://osf.io/tv57s> (<https://osf.io/tv57s>).

### Participants

50 participants for the main task were recruited on Prolific and were
paid \$10/hour. 8 subjects, 4 from the attractiveness task group and 4 from
the roundness task group, were excluded for incorrect validations.
<!-- they were able to proceed despite bad validation?--> After this
data exclusion, we ended up with 21 participants each for the
attractiveness task and the roundness task. The original sample size in
Shimojo et al. (2003) was 10 participants total.

### Procedure and Design

At the beginning of the experimental task, participants completed a
9-point eye-tracker calibration (each point appeared 3 times in random
order) and 3-point validation. The validation point appeared once at
center, middle left, and middle right locations in random order (see Figure\ \@ref(fig:E5-calibration-figure)).
<!-- does 9 points here mean that there were actually only 3 calibration points but each was shown 3 times so 9 total or is it 9 points shown 3 times so 27 total? If the latter, why is the coverage denser than for Experiment 4? -->
```{r E5-calibration-figure, child='group-e/E5-calibration-figure.Rmd'}

```

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure\ \@ref(fig:E5-example-trial)). Participants were randomly assigned to one of two tasks:
attractiveness or shape judgment. In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the "a" key on their keyboard to
select the face on the left and the "d" key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials (???). The order of
the 19 face pairs was random for each participant.

```{r E5-example-trial, fig.cap='An example of a critical trial from Experiment 5. (Text did not appear on each screen.)'}
knitr::include_graphics("group-e/E5-example-figure.jpeg")
```

### Materials and Norming

The faces in our replication were selected from a set of 1,000 faces
within the Flickr-Faces-HQ Dataset. (The face images used in Shimojo et
al. were from the Ekman face database and the AR face database.) These
images were chosen because the person in each image was looking at the
camera with a fairly neutral facial expression and appeared to be over
the age of 18. 27 participants were recruited on Prolific to participate
in stimulus norming (for attractiveness). They were paid \$XX for
completing the experiment. Data from 3 participants was excluded because
their mode response made up more than 50% of their total responses
<!-- is this a cutoff from Shimojo?-->, for a total of 24 participants
in the norming. They each viewed all 172 faces and were asked to rate
them on a scale from 1 (less attractive) to 7 (more attractive) using a
slider. Faces were presented one at a time and in a random order for
each participant. Following Shimojo et al., 19 face pairs were made by
matching two faces that had a difference in mean attractiveness ratings
that was 0.25 points or lower and that matched in gender, race, and age
group (young adult, adult, or older adult).

### Data analysis

In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position was then determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz.[TODO - CONFIRM]

## Results

```{r E5-analysis-replication, child='group-e/E5-analysis-replication.Rmd'}

```

### Calibration

```{r E5-analysis-calibration, child='group-e/E5-analysis-calibration.Rmd'}

```

### Effects of ROIs

```{r E5-analysis-roi, child='group-e/E5-analysis-roi.Rmd'}

```

## Discussion

```{=html}
<!-- posner cueing task
# Experiment 6

The sixth study was a replication attempt of Posner et al. ??.

## Methods

### Participants

### Procedure and Design

### Data analysis

## Results

### Replication

### Calibration

### ROIs

### Item Numbers

## Discussion
 -->
```
# Combined Analyses?

-   Pooling data from all experiments we can look at patterns in the
    calibration and validation data

# General Discussion

 - E1:

 - E2: 
    - replication of key result with 4 quadrants
    - calib quality doesn't seem to matter 
    - if attempting to control where people are looking, think about modifying task...

 - E3:

 - E4:

 - E5:

```{=html}
<!-- NOTES FROM EMAILS:

JOSH DE LEEUW: For calibration quality, there are a few ways you could go about it.

As a preliminary step filtering the data to trial_type == 'webgazer-validate' should get you a data frame with just the validation trials. We ran a validation trial immediately after calibration. The trial involves showing some set of dots on the screen for a few seconds each and recording all of the samples from webgazer for each dot. Each experiment defined a different set of dots for validation, depending on where the relevant ROIs were.

The data for each of these trials includes columns for:

raw_gaze: all the raw data points recorded from webgazer, organized by which point was shown on the screen
percent_in_roi:  The jsPsych plugin also calculates the percentage of samples for each dot that are within some specified radius. I think we used 150px for all of the experiments, but I can verify that.
average_offset: The average x,y offset of the cloud of points measured for each validation point. Also includes a value r, which is the radius of the circle centered at the midpoint of the cloud that includes half the points. Basically a measure of variance.
validation_points: coordinates of each of the points on the screen, in the same order as the other columns so that you could see, e.g., if calibration tends to be better on more central points.

We also included a step where if the validation was poor after the first attempt we repeated the calibration and validation procedure one additional time. So some subjects have more validation trials than others. If you look at my R Notebook for group A (just pushed to GH now) you can see some initial attempts to extract validation info and use it to exclude subjects. -->
```
\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```