### 2023-08-24 group-a/E1-analysis-comparison-to-JMW23.Rmd

---
title: "E1-analysis-comparison-to-JMW23"
author: "Ariel James"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)

knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = FALSE)
```

```{r import JMW23}
JMW1_info <- read_csv("group-a/JMW23-data/fyptrialinfo.csv")

JMW1_latency <- read.delim("group-a/JMW23-data/FYPAKFixationLatenciesPOSTVERB.txt", header = FALSE, sep = ",") #I'll come back to this

JMW1_data <- read.delim("group-a/JMW23-data/FYPTimecourse_by_subj.txt", header = FALSE, sep = ",")

#Add meaningful header row
tc_header = c("TARGET","AGENT", "COMP1", "COMP2", "COMP3", 
              "NON1", "NON2", "NON3", "NON4", "NON5",
              "ORDER", "SUBJ", "BIN", "COND", "AID") #AID is a unique item number

colnames(JMW1_data) <- tc_header

#Recode condition
JMW1_data$COND <- as.factor(recode(JMW1_data$COND, "0" = "filler", 
                                "1" = "pred", "2" = "neut"))
JMW1_info$Condition <- as.factor(recode(JMW1_info$Condition, "0" = "filler", 
                                "1" = "pred", "2" = "neut"))

#Remove fillers
JMW1_data_crit <- JMW1_data %>%
  filter(COND != "filler")
JMW1_data_crit <- droplevels(JMW1_data_crit)

#Pull out useful item info
JMW1_info$Non.relative <- JMW1_info$Non-JMW1_info$Von

```

```{r reshape data for comparison}
JMW.long <- JMW1_data_crit %>%
  select(!NON2:ORDER) %>%
  group_by(SUBJ, AID, BIN) %>%
  gather(key = "obj_type", value = "duration", TARGET, AGENT, COMP1, COMP2, COMP3, NON1) %>%
  #create "obj_class" col: name everything that isn't a targeT or agenT "DISTRACTOR" 
  mutate(obj_class = ifelse(str_ends(obj_type, "T"), obj_type, "DISTRACTOR")) %>%
  #create binary "is.fixating" variable
  mutate(is.fixating = ifelse(duration > 0, TRUE, FALSE)) %>%
  filter(BIN >= 0)
```

```{r compute cumulative fix}
#Write function (i.e. copy/paste function Josh deLeeuw wrote)
cumulative.fixation.calculation <- function(df){
  time <- seq(0,1950,50)
  fixations <- logical(length(time))
  hit.flag <- FALSE
  for(i in 1:length(fixations)){
    if(hit.flag){
      fixations[i] <- TRUE
    } else {
      val <- df %>% filter(BIN <= time[i]) %>% pull(is.fixating) %>% any()
      if(val == TRUE) {
        hit.flag <- TRUE
      }
      fixations[i] <- val
    }
  }
  out <- tibble(t.window = time, has.fixated = fixations)
  return(out)
}

#Now use it to summarize timecourse data
if(file.exists("group-a/JMW23-data/generated/JMW_cumulative.csv")){
  JMW_cumu<- read_csv("group-a/JMW23-data/generated/JMW_cumulative.csv")
} else {

  JMW_cumu <-  JMW.long%>%
    group_by(SUBJ, COND, AID, obj_type, obj_class) %>%
    summarize(cumulative.fixation.calculation(cur_data()))

  write_csv(JMW_cumu, file="group-a/JMW23-data/generated/JMW_cumulative.csv")
}
```

<!-- Then group by object class to calculate fixation probability for each kind of object, normalizing by the number of objects of that type. -->

```{r group cumu fix by object, message=FALSE, warning=FALSE}
JMW_cumu.windows.grouped <- JMW_cumu %>%
  group_by(SUBJ, COND, AID, obj_class, t.window) %>%
  summarize(cumu.p = mean(has.fixated)) 
```

<!-- Then collapse over trials to get a cumulative probability for the whole window for each subject -->

```{r collapse cumu fix over trials, message=FALSE, warning=FALSE}
JMW_cumu.windows.grouped.average <- JMW_cumu.windows.grouped %>%
  group_by(SUBJ, COND, obj_class, t.window) %>%
  summarize(cumu.p.avg = mean(cumu.p))
```

<!-- Collapse across subjects to generate the equivalent of the figure from Altmann & Kamide. -->

```{r create summary data for spag plot, message=FALSE, warning=FALSE}
JMW_cumu.summary <- JMW_cumu.windows.grouped.average %>%
  filter(obj_class %in% c('TARGET', 'DISTRACTOR')) %>%
  group_by(t.window, COND, obj_class) %>%
  summarize(cumulative.fixation.m = mean(cumu.p.avg))
```
```{r JMW-spaghetti-fig, fig.cap = "Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions."}

#AJ: this is where you left off. Everything above should work :)
JMW_spag_fig <- ggplot(JMW_cumu.summary, aes(x=t.window, y=cumulative.fixation.m, fill=obj_class, shape=COND, group=interaction(COND, obj_class)))+
  scale_fill_manual(values=c("white", "black"))+
  scale_shape_manual(values=c(22,21))+
  geom_line()+
  geom_point(size=2)+
  theme_bw()+
  theme(panel.grid=element_blank())+
  scale_y_continuous(limits=c(0,1))+
  scale_x_continuous(expand=c(0.01,0.01))+
  labs(x="Time from Verb Onset (ms)", y="Probability", shape="Verb Type", fill="Object Type") +
  geom_vline(xintercept = mean(JMW1_info$Non.relative))
#The "Object type" part of the key doesn't actually show which color is which
```