### 2024-02-17 group-a/E1-calibration-figure.Rmd
fixes to get figures to display and cross-ref correctly

---
title: "Experiment 1: Calibration Figure"
output: pdf
---

```{r include=FALSE}
library(ggplot2)
library(dplyr)
```

The task began with a 9-point eye-tracker calibration
and validation (Figure\ \@ref(fig:E1-calibration-figure)). 
During the experiment, the participants were
simultaneously presented with a visual image and a corresponding audio
recording of a spoken sentence. Participants had to input a keyboard
response indicating "yes" or "no" as to whether the sentence they heard
was feasible given the visual image. There were two practice trials to
ensure that participants understood the instructions
before they undertook the main portion of the experiment. Participants'
reaction times, keyboard responses, and looks to objects in the scene
were recorded for each trial.

```{r include=FALSE}
# calib: 
    # //top
    # [15, 2],
    # [85, 2],
    # //bottom
    # [15, 95],
    # [85, 95],
    # 
    # //inner top
    # [32.5, 26],
    # [67.5, 26],
    # //inner bottom
    # [32.5, 72.5],
    # [67.5, 72.5],
    # 
    # [50, 50]
calibration_points <- matrix(c(
  c(15,2),
  c(85,2),
  c(15,95),
  c(85,95),
  c(32.5,26),
  c(67.5,26),
  c(32.5,72.5),
  c(67.5,72.5),
  c(50,50)
), ncol=2, byrow=T)
# valid: 
#      [,1] [,2]
 # [1,] 15.0  2.0
 # [2,] 85.0  2.0
 # [3,] 15.0 95.0
 # [4,] 85.0 95.0
 # [5,] 22.5 26.0
 # [6,] 67.5 26.0
 # [7,] 22.5 72.5
 # [8,] 67.5 72.5
 # [9,] 50.0 50.0
validation_points <- matrix(c(
  c(15,2),
  c(85,2),
  c(15,95),
  c(85,95),
  c(22.5, 26),
  c(67.5, 26),
  c(22.5, 72.5),
  c(67.5, 72.5),
  c(50,50)
), ncol=2, byrow=T)

colnames(calibration_points) <- c("x","y")
colnames(validation_points) <- c("x","y")

df <- as_tibble(calibration_points) %>% bind_rows(as_tibble(validation_points))
df$point <- c(rep("calibration", nrow(calibration_points)), rep("validation", nrow(validation_points)))
```

```{r E1-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 1. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```