### 2024-02-17 group-c/E3-calibration-figure.Rmd

---
title: "Experiment 3: Calibration Figure"
output: html_notebook
---

```{r include=FALSE}
library(ggplot2)
library(dplyr)
```

The task began with a 7-point eye-tracker calibration (each point was
presented 3 times in a random order) and validation with 3 points (each
presented once). The point locations were designed to focus calibration
on the center of the screen and the middle of the left and right halves
of the screen (Figure\ \@ref(fig:E3-calibration-figure)).

```{r include=FALSE}
calibration_points <- matrix(c(
  c(10,50),
  c(50,50),
  c(90,50),
  c(25,25),
  c(25,75),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)

validation_points <- matrix(c(
  c(10,50),
  c(50,50),
  c(90,50)
), ncol=2, byrow=T)

colnames(calibration_points) <- c("x","y")
colnames(validation_points) <- c("x","y")

df <- as_tibble(calibration_points) %>% bind_rows(as_tibble(validation_points))
df$point <- c(rep("calibration", nrow(calibration_points)), rep("validation", nrow(validation_points)))
```

```{r E3-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 3. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```