### 2024-02-17 group-d/E4-calibration-figure.Rmd

---
title: "Experiment 4: Calibration Figure"
output: pdf
---

```{r include=FALSE}
library(ggplot2)
library(dplyr)
```

After the eye-tracking calibration and validation (Figure\ \@ref(fig:E4-calibration-figure)), participants went through an audio
test so they could adjust the audio on their computer to a comfortable
level. Before beginning the experiment, they were given instructions
that four objects would appear, an audio prompt would play, and they
should do their best to use their mouse to act out the instructions.
They then went through three practice trials which were followed by 54
critical trials and 24 filler trials presented in a random order.

```{r include=FALSE}
# calib: [25,25],[25,75],[50,50],[75,25],[75,75]
calibration_points <- matrix(c(
  c(25,25),
  c(25,75),
  c(50,50),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)
# valid: 
#     [,1] [,2]
# [1,]   25   25
# [2,]   25   75
# [3,]   50   50
# [4,]   75   25
# [5,]   75   75
validation_points <- matrix(c(
  c(25,25),
  c(25,75),
  c(50,50),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)

colnames(calibration_points) <- c("x","y")
colnames(validation_points) <- c("x","y")

df <- as_tibble(calibration_points) %>% bind_rows(as_tibble(validation_points))
df$point <- c(rep("calibration", nrow(calibration_points)), rep("validation", nrow(validation_points)))
```

```{r E4-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 4. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration.(In this experiment all the same locations were used for both calibration and validation.)", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```