### 2024-08-20 manuscript.Rmd
experiment 1 edits; split into 1a and 1b

---
title             : "What paradigms can webcam eye-tracking be used for? Attempted replications of five “classic” cognitive science experiments"
shorttitle        : "Webcam eye-tracking paradigms"

author:
  - name: Joshua R. de Leeuw
    affiliation: '1'
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Investigation
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    corresponding: yes
    email: jdeleeuw@vassar.edu
    address: 124 Raymond Ave, Poughkeepsie, NY 12604, USA
  - name: Rachel Ryskin
    affiliation: '2'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: rryskin@ucmerced.edu
  - name: Ariel N. James
    affiliation: '3'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: ajames2@macalester.edu
  - name: Joshua K. Hartshorne
    affiliation: '4'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: joshua.hartshorne@bc.edu
  - name: Haylee Backs
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: hbacks@vassar.edu
  - name: Nandeeta Bala
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: nbala@vassar.edu
  - name: Laila Barcenas-Meade
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: lbarcenasmeade@vassar.edu
  - name: Samata Bhattarai
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: sbhattarai@vassar.edu
  - name: Tessa Charles
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: tcharles@vassar.edu
  - name: Gerasimos Copoulos
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: gcopoulos@vassar.edu
  - name: Claire Coss
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: ccoss@vassar.edu
  - name: Alexander Eisert
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: aeisert@vassar.edu
  - name: Elena Furuhashi
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: efuruhashi@vassar.edu
  - name: Keara Ginell
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: kginell@vassar.edu
  - name: Anna Guttman-McCabe
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: aguttmanmccabe@vassar.edu
  - name: Emma (Chaz) Harrison
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: eharrison@vassar.edu
  - name: Laura Hoban
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: lhoban@vassar.edu
  - name: William A. Hwang
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: whwang@vassar.edu
  - name: Claire Iannetta
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: ciannetta@vassar.edu
  - name: Kristen M. Koenig
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: kkoenig@vassar.edu
  - name: Chauncey Lo
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: chaunceylo@vassar.edu
  - name: Victoria Palone
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: vpalone@vassar.edu
  - name: Gina Pepitone
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: gpepitone@vassar.edu
  - name: Margaret Ritzau
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: mritzau@vassar.edu
  - name: Yi Hua Sung
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: ysung@vassar.edu
  - name: Lauren Thompson
    affiliation: '1'
    role:
      - Investigation
      - Methodology
      - Software
    email: lthompson@vassar.edu

affiliation:
  - id: '1'
    institution: Cognitive Science Department, Vassar College
  - id: '2'
    institution: Department of Cognitive & Information Science, University of California,
      Merced
  - id: '3'
    institution: Psychology Department, Macalester College
  - id: '4'
    institution: Department of Psychology & Neuroscience, Boston College

abstract: |
  Web-based data collection allows researchers to recruit large and diverse samples with fewer resources than lab-based studies require. Recent innovations have expanded the set of methodolgies that are possible online, but ongoing work is needed to test the suitability of web-based tools for various research paradigms. Here, we focus on webcam-based eye-tracking; we tested whether the results of five different eye-tracking experiments in the cognitive psychology literature would replicate in a webcam-based format. Specifically, we carried out five experiments by integrating two javascript-based tools: js.psych and a modified version of Webgazer.js. In order to represent a wide range of applications of eye-tracking to cognitive psychology, we chose two psycholinguistic experiments, two memory experiments, and a decision-making experiment. These studies also varied in the type of eye-tracking display, including screens split into halves (Exps. 3 and 5) or quadrants (Exps. 2 and 4), or composed scenes with regions of interest that varied in size (Exp. 1). Outcomes were mixed. The least successful replication attempt was Exp. 1; we did not obtain a condition effect in our remote sample (1a), nor in an in-lab follow-up (1b). However, the other four experiments were more successful, replicating a blank-screen effect (Exp. 2), a novelty preference (Exp. 3), a verb bias effect (Exp. 4), and a gaze-bias effect in decision-making (Exp. 5). These results suggest that webcam-based eye tracking can be used to detect a variety of cognitive phenomena, including those with sensitive time, although paradigms that require high spatial resolution should be adapted to coarser quadrant or split-half displays.

keywords          : "eye-tracking, online, webcam, jsPsych, cognitive science"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
r_refs("references.bib")

colorize <- function(x, color) { # for adding colored text (e.g., for notes to self)
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

The use of eye-tracking to study cognition took off when Alfred Yarbus used suction cups to affix a mirror system to the sclera of the eye in order to monitor eye position during the perception of images [@yarbus1967eye]. In one study, participants viewed a painting depicting multiple people in a complex interaction inside of a 19th century Russian home. Yarbus showed, among other things, that the scan paths and locations of fixations were largely dependent on the instructions given to participants (e.g., View the picture freely vs. Remember the position of the people and objects in the room). In other words, the cognitive processing that the individual is engaged in drives the visuo-motor system. Since these findings, eye-tracking has become a central method in cognitive science research [for reviews, see @hayhoe2005eye, @richardsonEyeTrackingResearch2004, @rayner1998eye]. For example, gaze location during natural scene perception is used to test theories of visual attention [e.g., @henderson2017meaning], and eye-movements during auditory language comprehension, using the "visual world paradigm," demonstrated the context-dependent and incremental nature of language processing [e.g., @tanenhaus1995].

An important limitation of the eye-tracking methodology is that it has typically required costly equipment (eye-trackers can range in price from a few thousand dollars to tens of thousands of dollars), particular laboratory conditions (a quiet room with consistent indoor lighting conditions), and a substantial time investment (e.g., bringing participants into a laboratory one at a time). This limits who can conduct eye-tracking research -- not all researchers have the necessary resources -- and who can participate in eye-tracking research. Most eye-tracking study participants are from western, educated, industrialized, rich, and democratic [WEIRD; @henrich2010weirdest] convenience samples [but see @ryskin2023real], which diminishes the generalizability of the findings and the scope of conclusions that can be drawn about human cognition. Likewise, the sample sizes for in-lab experiments are usually orders of magnitude smaller than what statisticians recommend [@nosek2022replicability].

A robust solution to all these problems is online experiments, particularly with volunteer citizen scientists as participants [@li2024developmental; @hartshorne2019thousand; @reinecke2015labinthewild; @gosling2010wired]. Historically, this option has not been available for eye-tracking. This began to shift with the widespread incorporation of cameras into computers; researchers have long used frame-by-frame analysis of video for low-resolution eyetracking [e.g., @snedeker2003using]. Unfortunately, this can be extremely time-intensive, making large-sample studies unrealistic. In recent years, image analysis has improved to the point where this work can be automated with reasonable accuracy [@papoutsaki2016webgazer; @zheng2018rapid; @burton2014comparison; @skovsgaard2011evaluation]. However, webcam-based eyetracking only started to be used regularly in research with the advent of `Webgazer.js` [@papoutsaki2016webgazer], a webcam-based Javascript plug-in that works in the browser and which can be integrated with any Javascript web interface, including `jsPsych` [@deleeuwJsPsychJavaScriptLibrary2015], `Gorilla` [@anwyl2020gorilla], or `lab.js` [@henninger2021lab].

Given the potential game-changing nature of webcam-based eye-tracking, a number of research groups have investigated how well it works. There are two potential limitations to webcam-based eye-tracking. First, the spatial and temporal resolution is less than what is achievable with an infrared system. Second, testing subjects over the Internet involves less control: subjects may be unable or unwilling to calibrate equipment, adjust lighting, etc., to the same level of precision typical of in-lab studies. Nearly all this work has used `Webgazer.js` --- most in combination with jsPsych, but some with Gorilla or hand-built integrations.

Results to date have been encouraging. @semmelmann2018online found data quality was reasonable for fixation location and saccades in fixation, smooth pursuit, and free-viewing tasks, though data collected online through a crowdsourcing platform was slightly more variable and timing was somewhat delayed compared to data collected in the lab. Several researchers successfully replicated well-known findings from the sentence-processing literature involving predictive looks [@degen2021seeing; @prystauka2023online; @vos2022comparing; @van2023validation]. @yang2021webcam successfully replicated a well-established link between value-based decision-making and eye gaze [see also @van2023validation].

While promising, there are some salient limitations. First, many of the studies report effects that are smaller or later than what had been previously observed in the lab [@slim2022moving; @degen2021seeing; @van2023validation]. Importantly, accurate timing on a web browser is not trivial [@passell2021cognitive; @de2016psychophysics], and subtle programming choices can significantly affect the accuracy of `WebGazer.js` timing [@yang2021webcam].^[See also discussion at https://github.com/jspsych/jsPsych/discussions/1892] Since most of the prior work did not address these timing issues, it is not clear how many of the reported lab/web differences would resolve.

Second, prior work has focused on studies with relatively coarse-grained regions of interest [but see @semmelmann2018online], dividing the screen either in half or in quadrants. This is particularly salient with [@prystauka2023online], which simplified [@altmannIncrementalInterpretationVerbs1999]'s design so that regions of interest are quadrants rather than the finer-grained ROIs used in the original. Certainly, webcam eyetracking will not be as spatially fine-grained as an infrared eyetracker, but we do not yet have a good sense of the limits.

Finally, the prior work has focused on a relatively limited range of methods. Different paradigms have different technical requirements and analyze the results differently. As a result, the breadth of utility of webcam eye-tracking is unclear.

# Present work

In order to validate the online eye-tracking methodology, with the particular configuration known to have the greatest temporal precision, `jsPsych` and a modification of `Webgazer`, we set out to reproduce five previously published studies representing a variety of questions, topics, and paradigms. The goal was to examine the strengths and weaknesses of webcam eye-tracking for common paradigms in cognitive science, across a broad range of research areas.

## Selection of Studies

Studies with large effect sizes and which are known to replicate are ideal targets for further replication; otherwise, it can be difficult to distinguish a failure of the method from a failure of the original study to replicate. In practice, replications (successful or otherwise) have only been reported for a small number of studies, so we ultimately included some studies with unknown replicability. We addressed this in several ways. First, replicating five very different studies from different research traditions decreases our reliance on any one study. Second, we include several ''sanity check'' analyses, such as the correlation between calibration accuracy and effect size. If the effect is real but there is noise from low-accuracy eye-tracking, this correlation should be substantial. Third, for two of the studies, we had comparison data collected in-lab either using `jsPsych` or a more traditional eyetracker technology, allowing us to directly assess the impact of differences in subject population and equipment/setting.

<!-- June 2022: shimojo: 964 citations, altmann: 1840, ryskin: 75, johansson 190, 127 -->
<!-- RR: switched ryskin to Snedeker & trueswell 2004 bcs my 75 looked sad -->

We chose five high-impact eye-tracking studies involving adult subjects [for an investigation of WebGazer's validity for developmental research, see @steffan2024validation for a comparison of remote WebGazer and in-lab anticipatory looking effects in 18-27-month-old participants]. Our goal was to include experiments from a range of topic areas (e.g., memory, decision making, psycholinguistics) and paradigms (two halves of the screen, visual world paradigm with four quadrants, visual world paradigm with "naturalistic" scenes). As noted above, we had a preference for well-established findings that are known to replicate, though for sake of diversity this was not always possible. Table \@ref(tab:studies-table) provides an overview of the five studies we selected.

```{r studies-table, echo=FALSE}
studies_table <- data.frame(
  Citation = c(
    "Altmann & Kamide, 1999",
    "Johansson & Johansson, 2014",
    "Manns, Stark, & Squire, 2000",
    "Snedeker & Trueswell, 2004",
    "Shimojo et al., 2003"
  ),
  `Topic Area` = c(
    "Psycholinguistics",
    "Memory",
    "Memory",
    "Psycholinguistics",
    "Decision Making"
  ),
  `Paradigm` = c(
    "Natural Scenes",
    "Four Quadrants",
    "Two Halves",
    "Four Quadrants",
    "Two Halves"
  ),
  `Citations` = c(
    2130,
    259,
    134,
    487,
    1146
  ),
  check.names = FALSE
)

apa_table(
  studies_table,
  caption = "Studies selected for replication attempts. Citation counts based on Google Scholar (May 2024).",
  font_size = "small",
  digits = 0
)

```

# General Methods

## Participants

Participants completed the experiment remotely and were recruited
through the Prolific platform. In order to have access to the
experiment, participants had to meet the following criteria: 18 years of
age or older, fluency in English, and access to a webcam. All
participants provided informed consent. The online studies were approved by the
Vassar College Institutional Review Board.

In addition, an in-lab replication was conducted for Experiment 1. 
Information about the sample is given in the Experiment 1 Method sections. This study was approved by the Institutional Review Board at Boston College.

In order to have adequate statistical power and precision, we aimed for
2.5x the sample size of the original experiment, following the heuristic
of Simonsohn [@simonsohn2015]. In Experiment 5, the original sample size was so
small that we opted to collect 5x the number of participants to increase
precision. Because of budget and time constraints we were unable to
replace the data for subjects who were excluded or whose data was
missing due to technical failures.

## Equipment

We used a fork of the `webgazer.js` library for webcam eye-tracking
[@papoutsaki2016webgazer], implemented in `jsPsych`, a Javascript
library for running behavioral experiments in a web browser
[@deleeuwJsPsychJavaScriptLibrary2015]. Our fork included changes to
`webgazer.js` in order to improve data quality for experiments in which
the precise timing of stimulus onsets is relevant. Specifically, we
implemented a polling mode so that gaze predictions could be requested
at a regular interval, which improved the sampling rate considerably in
informal testing. This modification is similar to what Yang and Krajbich
[-@yang2021webcam] reported improved the sampling rate in their study of
webgazer. We also adjusted the mechanism for recording time stamps of
each gaze prediction, so that the time stamp reported by webgazer is
based on when the video frame is received and not when the computation
of the gaze point is finished.

## Eye-tracking Calibration and Validation

<!--took paragraph below from group D's methods -->

When participants began the experiment, they were notified the webcam
would be used for eye tracking but no video would be saved. They were
asked to remove glasses if possible, close any other tabs or apps, turn
off notifications, and make sure their face was lit from the front. The
webcam's view of the participant popped up on the screen, and
participants were asked to center their face in the box and keep their
head still. The experiment window then expanded to full screen, and
participants began the eye-tracking calibration.

During the calibration, dots appeared on the screen one at a time in
different locations, and the participants had to fixate them and click
on each one. Once they clicked on a dot, it would disappear and a new
one would appear in a different location on the screen. The locations of
calibration dots were specific to each experiment (details below) and
appeared in the areas of the screen where the visual stimuli would
appear during the main task in order to ensure that eye movements were
accurately recorded in the relevant regions of interest. After the
calibration was completed, the validation began. Participants were asked
to go through the same steps as the calibration, except that they only
fixated the dots as they appeared in different locations on the screen.
If accuracy on the validation was too low (fewer than 50% of looks
landed within a 200 px radius of the validation points<!--check!-->),
participants were given an opportunity to re-start the calibration and
validation steps. If the second attempt also lead to low validation
accuracy, participants were informed that they could not participate in
the study.

## Pre-registration

These data were collected within the context of an undergraduate research methods course. Groups of students (co-authors) designed and programmed experiments in jsPsych, pre-registered their planned analyses, and collected data through Prolific under the supervision of the first author. The OSF repositories associated with these experiments are linked in the methods sections of each individual study. Note that in the current paper we expand on those pre-registered analyses (e.g., including analyses of the calibration quality). All analysis code underlying this paper can be found in the Github repository: https://github.com/jodeleeuw/219-2021-eyetracking-analysis

## Data Pre-processing

We used `r cite_r("references.bib")` for all our analyses.

# Experiment 1a

<!-- group a -->

The first study was a replication attempt of
@altmannIncrementalInterpretationVerbs1999. Altmann and Kamide used the
visual world eye-tracking paradigm [@tanenhaus1995] to show that
meanings of verbs rapidly constrain the set of potential subsequent
referents in sentence processing. For example, when looking at the
display in Figure\ \@ref(fig:E1-example-trial) and listening to a sentence like
"The boy will eat the...," participants are more likely to look at the cake than 
when they hear "The boy will move the...," in which case they tend to look at the
train, presumably because cakes are edible and trains are not. Semantic
information available at the verb is used to anticipate upcoming
linguistic input.

## Method

All stimuli, experiment scripts, data, analysis scripts, and a
pre-registration are available on the Open Science Framework at
<https://osf.io/s82kz>.

### Participants
Sixty participants were paid \$2.60 for their participation. Our sample size
of participants was determined by the total run time of our experiment,
\~10 minutes, and the allotted funding from the Vassar College Cognitive
Science Department. From this information, we calculated a reasonable
number of participants we could afford to compensate on Prolific. Note
that the sample size of the original study was 24. For unknown reasons,
2 of the subjects' results were not recorded, so in the analysis, we
worked with data collected from 58 participants.

### Procedure

```{r E1 calibration figure, child = 'group-a/E1-calibration-figure.Rmd' }

```

```{=html}
<!-- Qs: 
-keyboard keys? what was the exact phrasing of the instruction? 
-Any delay between when visual scene appears and the audio starts? 
-Does trial end as soon as participant responds?-->
```
### Materials and Design

The visual stimuli were created through Canva and depicted an agent
accompanied by four to five objects in the scene (see Figure\ \@ref(fig:E1-example-trial)). On
critical trials, participants heard one of two sentences associated with
the scene. In the restrictive condition, the sentence (e.g., "The boy
will eat the cake") contained a verb (e.g., "eat") which restricts the
set of possible subsequent referents (e.g., to edible things). Only the
target object (e.g., the cake) was semantically consistent with the
verb's meaning. In the non-restrictive condition, the sentence (e.g.,
"The boy will move the cake") contained a verb (e.g., "move") which does
not restrict the set of possible subsequent referents. The target object
(e.g., the cake) as well as the distractor objects (e.g., the train, the
ball, etc.) were semantically consistent with the verb's meaning. Both
sentences were compatible with the scene, such that the correct keyboard
response for the critical trials was "yes." Filler trials consisted of
scenes that also contained an agent surrounded by objects as in the critical trials, but corresponding sentences named an object that was not present in the scene. The correct keyboard response for the filler
trials was "no."

```{r E1-example-trial, fig.cap='Example trial from Experiment 1. Participants would hear a sentence (e.g., "The boy will eat the cake") and respond according to whether the sentence matched the picture.'}
knitr::include_graphics("group-a/E1-example-trial.jpeg")
```

Each participant was presented with 16 critical trials (eight in
the restrictive condition, eight in the non-restrictive condition) and
16 fillers for a total of 32 trials. The order of trials and the
assignment of critical scene to condition was random on a
subject-by-subject basis.<!-- right?-->

## Results
Looks to the objects in the scene were time-locked to the onset of the
verb, the offset of the verb, onset of the post-verbal determiner, and
onset of the target noun. ROIs were defined by creating boxes around each object in the scene. The size of each box was determined by taking the height and width of the given object and adding 20 pixels of padding. Each scene contained an agent region, a target region, and three or four distractor regions.

```{r E1-analysis-replication, child = 'group-a/E1-analysis-replication.Rmd'}

```

## Discussion
Across three different tests of the hypothesis that listeners will use verb semantics to anticipate the upcoming referent, we found results that were numerically consistent with the hypothesis but not statistically significant. A comparison to published data [@james2023language] demonstrates that looks to the objects in the scene (relative to background or off-screen looks) were depressed across conditions and objects. After eliminating participants with validation accuracy under 50% and/or 10% or fewer fixations to any ROIs, we were left with only 22 of the original 60 participants. Analyses on this subset resulted in one effect reaching statistical significance (shorter target fixation latencies in the restricting vs. non-restricting verb condition). 
Given that nearly two-thirds of participants had poor data quality, we ran a follow-up webcam study in a lab setting order to isolate the potential causes.

# Experiment 1b
Experiment 1b tested whether the failure to replicate significant condition effects in Experiment 1a was due to features of conducting the study remotely (i.e. varied experimental settings and apparatuses, lower compliance) rather than webcam-based eye-tracking or `webgazer` _per se_. Thus, Experiment 1b took place in a lab setting with undergraduate participants but otherwise used the same Method as Experiment 1a.

## Method

### Participants
Forty-nine participants completed the study in a lab setting. They were recruited via the Boston College subject pool. Participants needed to be 18 years of age or older and native speakers of English.

### Procedure
After being greeted by the experimenter and completing the informed consent form, participants followed on-screen prompts to complete the study, including calibration, as described in the Experiment 1a Procedure.

### Apparatus
HELP

### Materials and Design
Materials were identical to those in Experiment 1a.

## Results
```{r E1-analysis-JoshH-lab, child = 'E1-analysis-comparison-to-Hartshorne.Rmd'}

```

## Discussion
Overall, the results of Experiment 1 paint a sobering picture of web-based eye-tracking. In Experiment 1a, results from the remote samplewere in the expected direction but effects were smaller and delayed relative to previous work and failed to reach statistical significance. To test whether this could be explained by the variability in experimental settings across participants, we replicated our procedure in a lab setting. Surprisingly, the results in the in-lab study were less aligned with previous work; critical effects were not significant nor were they in the expected direction numerically. This suggests the problem was not with running the study online or due to working with a more diverse population via Prolific.

Taken together, the instability of these effects might suggest that this paradigm is not well-suited for webcam-based eye-tracking. Notably, the ROIs were tightly drawn around the five to six objects in each scene (drawing larger ROIs in these scenes would have led to overlapping objects) and thus, analyses were unforgiving of inaccurate calibration. Further evidence comes from a recent `webgazer` study that successfully replicated a modified version of the Altmann and Kamide study with only four objects, each in separate quadrants, allowing for larger, more distinct ROIs [@prystauka2023online].

The next four experiments test paradigms with more generous ROIs.

# Experiment 2

<!-- group b -->

The second study was a replication attempt of @johanssonLookHereEye2014,
which examined how visuospatial information is integrated into memory
for objects. They found that, during memory retrieval, learners
spontaneously look to blank screen locations where pictures were located
during encoding [see @spiveyOculomotorMechanismsActivated2001] and that
this spatial reinstatement facilitates retrieval of the picture.

## Method

All stimuli, experiment scripts, data, analysis scripts, and a
pre-registration are available on the Open Science Framework at
<https://osf.io/xezfu/>.

### Participants

60 participants were paid for their participation.
The sample size was motivated in part by budget constraints, but was
nonetheless 2.5x larger than the original sample size of 24). 
Data from 1 participant were not properly
recorded due to unknown technical issues, so data from 59 participants
were included in all analyses to follow.

### Procedure

```{r E2 calibration figure, child='group-b/E2-calibration-figure.Rmd'}

```

The experiment consisted of two blocks each composed of an encoding
phase and a recall phase. During the encoding phase, participants saw a
grid indicating the four quadrants of the screen. Each quadrant
contained six images of items belonging to the same category (see Figure\ \@ref(fig:E2-example-trial)). 
The four categories were humanoids, household objects, animals, and
methods of transportation. 
Each of the four quadrants was presented one at a time. First, a list of
the items in the quadrant was shown, then the pictures of items were displayed in the quadrant. 
For each item, participants used their arrow keys to indicate whether the object was facing left or right. After the participant identified the direction of each
item, they would have an additional 30 seconds to encode the name and
orientation of each item in the quadrant. Finally, after all four quadrants
were presented, participants were shown the full grid of
24 items and had 60 seconds to further encode the name and orientation
of each item.

```{r E2-example-trial, fig.cap='Example trial from Experiment 2.'}
knitr::include_graphics("group-b/E2-example-figure.jpeg")
```

During the recall phase, participants listened to statements and
responded by pressing the 'F' key for false statements and 'T' for true
ones. Each statement fell into either an interobject or intraobject
condition. Interobject statements were those that compared two different
items in the grid (e.g. "The skeleton is to the left of the robot"),
while intraobject statements were those that asked about the orientation
of a single item (e.g. "The bus is facing right"). There were 48 total
statements, with 24 interobject and 24 intraobject statements split
evenly among the four quadrants. While listening to these statements, in
the free-viewing block, participants saw a blank screen and were allowed
to freely gaze around the screen. During the fixed-viewing block,
participants were asked to fixate a small cross in the center of the
screen throughout the recall phase. In both cases, the mouse was
obscured from the screen. Participants were randomly assigned to see the
fixed-viewing or free-viewing block first. Different images were used in each block.

After completing both encoding-recall blocks, participants were asked to
answer a few survey questions (such as whether they wore glasses or
encountered any distractions).

The primary methodological difference between this replication and
Johansson and Johansson's study was that the original study included two
additional viewing conditions that were omitted from this replication
due to time constraints. In those two conditions, participant were
prompted to look to a specific quadrant (rather than free viewing or
central fixation) which either matched or mismatched the original
location of the to-be-remembered item.

## Results

```{r E2-analysis-replication, child='group-b/E2-analysis-replication.Rmd'}

```

### Calibration

```{r E2-analysis-calibration, child='group-b/E2-analysis-calibration.Rmd'}

```

## Discussion

As in @johanssonLookHereEye2014 and @spiveyOculomotorMechanismsActivated2001, during memory retrieval, learners spontaneously look to blank screen locations where pictures were located
during encoding, suggesting that visuospatial information is integrated into the memory
for objects. However, we did not observe a memory benefit, in terms of speed or accuracy, of spatial reinstatement via gaze position during retrieval of the picture. We can speculate that this may be due to the fact that participants struggled to maintain their gaze fixed in the center in the fixed-viewing condition, such that the difference between the fixed- and free-viewing conditions was minimal. Crucially for the current purposes, the webcam-based eye-tracking measurements were successful in replicating the key eye-tracking results.

# Experiment 3

<!-- group c -->

The third study was a partial replication attempt of
@mannsVisualPairedcomparisonTask2000. This experiment used the visual
paired-comparison, which involves presenting a previously-viewed image
and novel image together and measuring the proportion of time spent
looking at each image. The expected pattern of results is that
participants will look more at novel objects. They
@mannsVisualPairedcomparisonTask2000 hypothesized that this pattern of
behavior could be used to measure the strength of memories. If a viewer
has a weak memory of the old image, then they may look at the old and
new images roughly the same amount of time. They tested this in two
ways. First, they showed participants a set of images, waited five
minutes, and then paired those images with novel images. They found that
participants spent more time (58.8% of total time) looking at the novel
images. They then measured memory performance one day later and found
that participants were more likely to recall images that they had spent
less time looking at during the visual paired-comparison task the
previous day.

## Method

The stimuli, experimental code, and data and analysis scripts can be
found on the Open Science Framework at <https://osf.io/k63b9/>.
The pre-registration for the study can be
found at <https://osf.io/48jsv> . We
inadvertently did not create a formal pre-registration using the OSF
registries tool, but this document contains the same information and is
time stamped prior to the start of data collection. <!-- Students failed
to pre-register this one and I missed it. There's an unofficial
pre-registration in the OSF site. -->

### Participants

Our pre-registered target was 50 participants. 51 participants completed
the first day of the experiment and 48 completed the second day.
Following Manns et al., we excluded 3 participants due to perfect
performance on the recognition memory test because this prevents
comparison of gaze data for recalled vs. non-recalled images. Our final
sample size was 45 participants.

### Procedure

```{r E3 calibration figure, child='group-c/E3-calibration-figure.Rmd'}

```

The experiment was administered over the course of two
consecutive days. It consisted of three sections: a presentation phase,
a test phase, and a recognition test. The first two phases occurred on
the first day, while the recognition test occurred on the second day.

During the presentation phase, participants viewed 24 pairs of identical
color photographs depicting common objects. Each pair was presented for
5 seconds and an interval of 5 seconds elapsed before the next pair was
shown. The order of the photographs was randomized and different for
each participant. After completion of the presentation phase,
participants were given a 5-minute break during which they could look
away from the screen.

After the break, they were prompted to complete the eye-tracking
calibration again before beginning the test phase. During this phase,
participants again viewed 24 pairs of photographs with an interstimulus
duration of 5 seconds. In each pair, one photograph was previously seen
during the presentation phase, while the other was new. Which pictures
were old or new was counterbalanced across participants.<!-- 2 lists?-->
For half of the participants in each counterbalancing group, the new and
old photographs were
reversed.<!-- not sure what that last sentence means: is it about the side of the screen? or still just describing that there were 2 lists-->

Approximately 24 hours after completing the first session, with a leeway
interval of 12 hours to accommodate busy schedules, participants were
given the recognition test. It consisted of 48 photographs, presented
one at a time. Each was shown on the screen for 1 second, followed by a
1 second interstimulus interval. Half of the photographs had been
viewed twice on the previous day and were deemed the "targets." The
other half depicted an object with the same name as an object in one of
the old photographs, but had not been viewed before, deemed "foils."
Each photograph remained on the screen until the participants indicated
whether or not they had seen it before by pressing 'y' for yes and 'n'
for no. After they pressed one of the two keys, a prompt on the screen
asked them to rate their confidence in their answer from 1 as a "pure
guess" to 5 as "very sure." by clicking on the corresponding number on
the screen. No feedback on their responses was given during the test.

The experimental design is visually depicted in Figure\ \@ref(fig:E3-design-schematic).

There were two modifications we made to the methods of the original
experiment. As we were only replicating the declarative memory component
of the original experiment, we did not have a "priming group."
Therefore, we followed only the procedure for the "looking group."
Additionally, for each section of the study, the stimuli were presented
on a single screen instead of two screens due to the constraints of the
online experiment format.

```{r E3-design-schematic, fig.cap='Schematic of the design of Experiment 3'}
knitr::include_graphics("group-c/E3-example-figure.jpeg")
```

### Materials

`r colorize('Images were selected XXX...','red')`

## Results

```{r E3-analysis-replication, child='group-c/E3-analysis-replication.Rmd'}

```

### Effects of ROIs

```{r E3-analysis-ROIs, child='group-c/E3-analysis-roi.Rmd'}

```

### Calibration

```{r E3-analysis-calibration, child='group-c/E3-analysis-calibration.Rmd'}

```

## Discussion

As in @mannsVisualPairedcomparisonTask2000, participants looked more at novel images than previously seen images. This effect was consistent for ROIs based on the images and for the coarser ROIs based on two halves of the display. A day later, participants were also able to discriminate the images they had seen from foil images they had not seen during the previous session. However, there was no evidence that memory performance on day 2 was related to looking time on day 1. Calibration quality did not appear to impact this relationship.

# Experiment 4

<!-- group d -->

The fourth study was a replication attempt of Experiment 1 in
@ryskinVerbBiasesAre2017, which was closely modeled on
@snedekerDevelopingConstraintsParsing2004a. These studies used the
visual world paradigm to show that listeners use knowledge of the
co-occurrence statistics of verbs and syntactic structures to resolve
ambiguity. For example, in a sentence like "Feel the frog with the
feather," the phrase "with the feather" could be describing the frog, or
it could be describing the instrument that should be used to do the
"feeling." When both options (a frog holding a feather and a feather by
itself) are available in the visual display, listeners rely on the
verb's "bias" (statistical co-occurrence either in norming or corpora)
to rapidly choose an action while the sentence is unfolding. .

## Method

The stimuli, experimental code, and data and original analysis scripts can be
found on the Open Science Framework at the following link,
<https://osf.io/x3c49/>. The pre-registration
for the study can be found at <https://osf.io/3v4pg>.

### Participants

57 participants were paid \$2.50 for their participation. A sample size of 60 was initially chosen (but not reached in time) because we wanted to replicate the experiment with greater statistical power. Note that the original study had a sample size of 24.

### Procedure

```{r E4 calibration figure, child='group-d/E4-calibration-figure.Rmd'}

```

During a trial, four pictures were displayed (target animal, target
instrument, distractor animal, distractor instrument), one in each
corner of the screen, and participants heard an audio prompt that
contained instructions about the action they needed to act out (e.g.,
"Rub the butterfly with the crayon"; see Figure\ \@ref(fig:E4-example-trial))[^1]. Using their
cursor, participants could act out the instructions by clicking on
objects and moving them or motioning over the objects[^2]. After the
action was completed, the participants were instructed to press the
space bar which led to a screen that said "Click Here" in the middle in
order to remove bias in the eye and mouse movements from the previous
trial. The experiment only allowed the participants to move on to the
next trial once the audio was completely done playing and the mouse had
been moved over at least one object.

[^1]: In the original study, the pictures appeared one by one on the
    screen and their names were played as they appeared. We removed this
    introductory portion of the trial to save time

[^2]:  As opposed to the original study we recorded mouse movement
    instead of clicking behavior since not all of the audio prompts
    required clicking. For example, the sentence "locate the camel with
    the straw" may not involve any clicking but rather only mousing over
    the camel.

```{r E4-example-trial, fig.cap='An example of a critical trial from Experiment 4 for the sentence “Rub the butterfly with the crayon.” The butterfly is the target animal, the panda is the distractor animal, the crayon is the target instrument, and the violin is the distractor instrument.'}
knitr::include_graphics("group-d/E4-example-figure.jpeg")
```

### Materials

The images and audios presented to the participants were the same
stimuli used in the original study (available here <link>). The critical
trials were divided into modifier-biased, instrument-biased, and
equibiased conditions, and the filler trials did not contain ambiguous
instructions. Two lists of critical trials were made with different verb
and instrument combinations (e.g., "rub" could be paired with "panda"
and "crayon" in one list and "panda" and "violin" in the second list).
Within each list, the same verb was presented twice but each time with a
different target instrument and animal. The lists were randomly assigned
to the participants to make sure the effects were not caused by the
properties of the animal or instrument images used. The list of verbs
used can be found in Appendix A of the original study.

## Results

```{r E4-analysis-replication, child='group-d/E4-analysis-replication.Rmd'}

```

### Comparison to in-lab data

```{r E4-analysis-comparison-to-original-study, child='group-d/E4-analysis-comparison-to-original.Rmd'}

```

### Calibration

```{r E4-analysis-calibration, child='group-d/E4-analysis-calibration.Rmd'}

```

### Effects of ROIs

```{r E4-analysis-ROIs, child='group-d/E4-analysis-ROIs.Rmd'}

```

## Discussion

As in @ryskinVerbBiasesAre2017 and @snedekerDevelopingConstraintsParsing2004a, listeners' gaze patterns during sentences with globally ambiguous syntactic interpretations differed depending on the bias of the verb (i.e., modifier, instrument or equi). For modifier-biased verbs, participants looked more quickly at the target animal and less at the potential instrument than for instrument-biased verbs (and equi-biased verbs elicited a gaze pattern between these extremes). This pattern was stronger for those who achieved higher calibration accuracy and when quadrant-based ROIs were used compared to image-based ROIs.

# Experiment 5

<!-- group e -->

The fifth study was a replication attempt of @shimojoGazeBiasBoth2003,
which found that human gaze is actively involved in preference
formation. Separate sets of participants were shown pairs of human faces
and asked either to choose which one they found more attractive or which
they felt was rounder. Prior to making their explicit selection,
participants were increasingly likely to be fixating the face they
ultimately chose, though this effect was significantly weaker for
roundness discrimination.

Note that Shimojo and colleagues compare five conditions, of which we
replicate only the two that figure most prominently in their
conclusions: the "face-attractiveness-difficult task" and the
"face-roundness task".

## Method

All stimuli, experiment scripts, data, and analysis scripts are
available on the Open Science Framework at [https://osf.io/eubsc/](https://osf.io/eubsc/). 
The study pre-registration is available at
[https://osf.io/tv57s](https://osf.io/tv57s).

### Participants

50 participants for the main task were recruited on Prolific and were
paid \$10/hour. 8 subjects, 4 from the attractiveness task group and 4 from
the roundness task group, were excluded for incorrect validations.
<!-- they were able to proceed despite bad validation?--> After this
data exclusion, we ended up with 21 participants each for the
attractiveness task and the roundness task. The original sample size in
Shimojo et al. (2003) was 10 participants total.

### Procedure and Design

```{r E5 calibration figure, child='group-e/E5-calibration-figure.Rmd'}

```

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure\ \@ref(fig:E5-example-trial)). Participants were randomly assigned to one of two tasks:
attractiveness or shape judgment. In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the "a" key on their keyboard to
select the face on the left and the "d" key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials. The order of
the 19 face pairs was random for each participant.

```{r E5-example-trial, fig.cap='An example of a critical trial from Experiment 5. (Text did not appear on each screen.)'}
knitr::include_graphics("group-e/E5-example-figure.jpeg")
```

### Materials and Norming

The faces in our replication were selected from a set of 1,000 faces
within the Flickr-Faces-HQ Dataset. (The face images used in Shimojo et
al. were from the Ekman face database and the AR face database.) These
images were chosen because the person in each image was looking at the
camera with a fairly neutral facial expression and appeared to be over
the age of 18. 27 participants were recruited on Prolific to participate
in stimulus norming (for attractiveness). They each viewed all 172 faces and were asked to rate
them on a scale from 1 (less attractive) to 7 (more attractive) using a
slider. Faces were presented one at a time and in a random order for
each participant. Data from 3 participants were excluded because
their mode response made up more than 50% of their total responses, for a total of 24 participants
in the norming.

Following Shimojo et al., 19 face pairs were selected by
identifying two faces that 1) had a difference in mean attractiveness ratings
that was 0.25 points or lower and 2) matched in gender, race, and age
group (young adult, adult, or older adult).

### Data analysis

In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position was then determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz.`r colorize("[TODO - CONFIRM]", 'red')`

## Results

```{r E5-analysis-replication, child='group-e/E5-analysis-replication.Rmd'}

```

### Calibration

```{r E5-analysis-calibration, child='group-e/E5-analysis-calibration.Rmd'}

```

### Effects of ROIs

```{r E5-analysis-roi, child='group-e/E5-analysis-roi.Rmd'}

```

## Discussion

Qualitatively, the results are similar to those of Shimojo et al., such that participants look more at the option that they ultimately choose. This gaze bias appears to be stronger for decisions about face attractiveness than shape, though this is not supported by the statistical analysis approach used in the original paper. The gaze patterns remained consistent for participants with better calibration accuracy.

```{=html}
<!-- posner cueing task
# Experiment 6

The sixth study was a replication attempt of Posner et al. ??.

## Methods

### Participants

### Procedure and Design

### Data analysis

## Results

### Replication

### Calibration

### ROIs

### Item Numbers

## Discussion

# Combined Analyses?

-   Pooling data from all experiments we can look at patterns in the
    calibration and validation data

 -->
```

# General Discussion

We conducted five attempted replication studies using different experimental paradigms from across the cognitive sciences. All were successfully implemented in `jsPsych` using the `webgazer` plugin, but replication success was mixed. Experiment 1 had the smallest ROIs due to the use of an integrated visual scene with five to six ROIs of varying size per scene, as opposed to ROIs corresponding to display halves or quadrants. Both attempts to replicate @altmannIncrementalInterpretationVerbs1999 were unsuccessful, despite the success of previous in-lab replications using infrared eye-tracking [e.g. @james2023language]. A previous conceptual replication of this paradigm using webcam-based eye-tracking [@prystauka2023online] was successful but used a four-quadrant visual world paradigm, rather than the "naturalistic" scenes used in the original study and in the current replication attempts. It is worth noting that removing variability related to participant environments (by conducting the webcam-tracking study in the lab) did not appear to improve the sensitivity of the paradigm. The primary limitation is likely to be the size of the ROIs.

Experiment 2 used the four quadrants of the participant's screen as ROIs. As in @johanssonLookHereEye2014 and @spiveyOculomotorMechanismsActivated2001, participants spontaneously looked to blank ROIs which previously contained to-be-remembered pictures. These results appeared to be robust to calibration quality. An additional manipulation, instructing participants to keep gaze fixed on a central point, was not successful. One possibility is that participants are less motivated to follow such instructions when an experimenter is not present in the same room with them. It may be possible to improve performance by emphasizing that this is an important aspect of the experiment or by providing additional training/practice in keeping the eyes still on one particular point.

Experiment 3 used two large ROIs (halves of the display in one analysis) and successfully replicated the novelty preference in terms of gaze duration shown in @mannsVisualPairedcomparisonTask2000. However, the subtler relationship between gaze duration and recognition memory on Day 2 was not replicated, despite the fact that participants were able to discriminate pictures they had seen from those they hadn't seen during that delayed test. Calibration quality did not appear to impact this relationship. More work is needed to understand whether delay manipulations can be practically combined with webcam eye-tracking.

Experiment 4 used the four quadrants of the participant's screen as ROIs. As in @ryskinVerbBiasesAre2017, listeners used knowledge of the co-occurrence statistics of verbs and syntactic structures to resolve ambiguous linguistic input ("Rub the frog with the mitten"). Across multiple time windows, participants looked more at potential instruments (mitten), when the verb (rub) was one that was more likely to be followed by a prepositional phrase describing an instrument with which to perform the action , as opposed to describing the recipient of the action (frog). Despite the qualitative replication of past findings, the overall rates of looks to various objects were much lower than in an in-lab study using infrared eye-tracking. This reduction may be related to measurement quality: effect sizes were greater for participants with higher calibration accuracy. Using the full quadrants as ROIs, rather than bounding boxes around the four images, also appeared to improve the measurement of the effect. Crucially, there was no evidence of a delay in the onset of effects relative to in-lab work, indicating that the modifications to `webgazer` that are made within the `jsPsych` plug-in successfully address the issues noted by @dijkgraaf2017predicting and suggesting that this methodology can be fruitfully used to investigate research questions related to the timecourse of processing.

Experiment 5, similar to Experiment 3, used two large ROIs (or halves of the display). As in @shimojoGazeBiasBoth2003 and in the recent webcam-based replication by @yang2021webcam, we saw that participants looked more at the face or shape that they ultimately chose during a judgment task. This gaze bias appears to be stronger for decisions about face attractiveness than shape, though this effect was not statistically significant.

In sum, the `webgazer` plug-in for `jsPsych` can be fruitfully used to conduct a variety of cognitive science experiments on the web, provided the limitations of the methodology are carefully considered. Studies with ROIs that take up half or a quarter of the participant's display, which encompasses a large number of common paradigms, are very likely to be successful, even when testing questions related to the timecourse of processing. However, the smaller the ROIs, the more important the calibration becomes. For instance, studies with four ROIs may want to exclude data from participants with less than 75% validation accuracy, whereas studies using two halves of the display as ROIs may not need to be so conservative. Studies with smaller ROIs (see Experiment 1) may not be appropriate for webcam eye-tracking in its current form.

```{=html}
<!-- NOTES FROM EMAILS:

JOSH DE LEEUW: For calibration quality, there are a few ways you could go about it.

As a preliminary step filtering the data to trial_type == 'webgazer-validate' should get you a data frame with just the validation trials. We ran a validation trial immediately after calibration. The trial involves showing some set of dots on the screen for a few seconds each and recording all of the samples from webgazer for each dot. Each experiment defined a different set of dots for validation, depending on where the relevant ROIs were.

The data for each of these trials includes columns for:

raw_gaze: all the raw data points recorded from webgazer, organized by which point was shown on the screen
percent_in_roi:  The jsPsych plugin also calculates the percentage of samples for each dot that are within some specified radius. I think we used 150px for all of the experiments, but I can verify that.
average_offset: The average x,y offset of the cloud of points measured for each validation point. Also includes a value r, which is the radius of the circle centered at the midpoint of the cloud that includes half the points. Basically a measure of variance.
validation_points: coordinates of each of the points on the screen, in the same order as the other columns so that you could see, e.g., if calibration tends to be better on more central points.

We also included a step where if the validation was poor after the first attempt we repeated the calibration and validation procedure one additional time. So some subjects have more validation trials than others. If you look at my R Notebook for group A (just pushed to GH now) you can see some initial attempts to extract validation info and use it to exclude subjects. -->
```
\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```