### 2024-11-02 group-c/E3-analysis-calibration.Rmd

---
title: "Experiment 3: Calibration Analysis"
output: html_notebook
---

```{r E3-cal load libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(jsonlite)
library(ggplot2)
```

```{r E3-cal load all data, include=FALSE}
data.files <- list.files('data', full.names = TRUE)
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  data.table$response <- as.character(data.table$response)
  return(data.table)
})
all.data <- bind_rows(data.tables)
```

```{r E3-cal Split into day 1 and day 2, include=FALSE}
study.id.day.1.group.1 <- "6062337f1a63a6788185d0e3"
study.id.day.1.group.2 <- "607de170bb44b7e25174421f"
study.id.day.2.group.1 <- "606233c0e42b980319b6a18e"
study.id.day.2.group.2 <- "607de2837874a824e2e8c6ec"

all.data <- all.data %>%
  mutate(day = if_else(study_id %in% c(study.id.day.1.group.1, study.id.day.1.group.2), 1, 2)) %>%
  mutate(group = if_else(study_id %in% c(study.id.day.1.group.1, study.id.day.2.group.1), 1, 2))

data.day.1 <- all.data %>% filter(day == 1)
data.day.2 <- all.data %>% filter(day == 2)
```

```{r E3-cal Select calibration data, include=FALSE}
eyetracking.calibration.data <- data.day.1 %>%
  filter(trial_type == "webgazer-validate") %>%
  select(subject_id, trial_index, phase,  percent_in_roi, average_offset) %>% 
  unnest(percent_in_roi)

eyetracking.calibration.summary <- eyetracking.calibration.data %>% 
  group_by(subject_id, trial_index, phase) %>% 
  summarize(mean_percent_in_roi = mean(percent_in_roi)) %>% 
  group_by(subject_id) %>% 
  mutate(calib_num = row_number())

subject.calibration.score <- eyetracking.calibration.summary %>%
  filter(phase=="boost-calibration") %>%
  select(subject_id, mean_percent_in_roi)
```

```{r E3-cal Select eyetracking data, include=FALSE}
eyetracking.data <- data.day.1 %>%
  filter(phase == "part2") %>%
  select(subject_id, webgazer_data, webgazer_targets, left_image_new, right_image_new, right_image, left_image) %>%
  tidyr::unpack(webgazer_targets) %>%
  tidyr::unpack(c("#left-img", "#right-img"), names_sep=".") %>%
  tidyr::unnest(webgazer_data)
```

```{r E3-cal Add ROI information, include=FALSE}
in.box <- function(x, y, left, right, top, bottom, padding){
  is.in.the.box <- x >= left - padding & x <= right + padding & y >= top - padding & y <= bottom + padding
  return(is.in.the.box)
}

eyetracking.data <- eyetracking.data %>%
  mutate(midpoint.x = `#left-img.right` + (`#right-img.left` - `#left-img.right`)/2) %>%
  mutate(in.left.half = x < midpoint.x) %>%
  mutate(in.right.half = x > midpoint.x)%>%
  mutate(in.target.half = if_else(left_image_new == TRUE, in.left.half, in.right.half))%>%
  mutate(in.distractor.half = if_else(left_image_new == TRUE, in.right.half, in.left.half)) %>%
  mutate(in.left.roi = in.box(x, y,`#left-img.left`, `#left-img.right`, `#left-img.top`, `#left-img.bottom`, 50))%>%
  mutate(in.right.roi = in.box(x, y,`#right-img.left`, `#right-img.right`, `#right-img.top`, `#right-img.bottom`, 50))%>%
  mutate(in.target.roi = if_else(left_image_new == TRUE, in.left.roi, in.right.roi))%>%
  mutate(in.distractor.roi = if_else(left_image_new == TRUE, in.right.roi, in.left.roi))
```

```{r E3-cal Calculate looking score, include=FALSE}
looking.score.data <- eyetracking.data %>% 
  group_by(subject_id, right_image, left_image, right_image_new, left_image_new) %>%
  summarise(looking.score.roi = sum(in.target.roi)/ sum(in.target.roi, in.distractor.roi),
            looking.score.half = sum(in.target.half) / sum(in.target.half, in.distractor.half)) %>%
  mutate(image = if_else(right_image_new, left_image, right_image))
```

```{r E3-cal Join looking score with calibration, include=FALSE}
merged.eyetracking <- looking.score.data %>%
  group_by(subject_id) %>%
  summarize(looking.score.M.roi = mean(looking.score.roi),
            looking.score.M.half = mean(looking.score.half))%>%
  left_join(subject.calibration.score, by=c("subject_id"))
```

<!--#### Calibration Accuracy-->

```{r E3-cal Plot Calibration Accuracy, echo=FALSE, eval=FALSE}
#This figure isn't referred to in the body of the text, and we have the scatterplots, so I'm cutting this for now
ggplot(subject.calibration.score, aes(x=mean_percent_in_roi))+
  stat_bin(binwidth = 10)+
  labs(x="Calibration Accuracy\n(% of samples within 200px of calibration point)")+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank())
```

<!--#### Correlation with Effects-->

To see if calibration success is correlated with the eye tracking effects, we calculated a calibration score for each participant. The calibration score was the average proportion of samples within 200 pixels of the validation points during the final validation phase before the eye tracking is performed. Calibration scores were not correlated with proportion looks to novel, regardless of whether scores were computed using ROIs (see Figure\ \@ref(fig:E3-cal-Plot-looking-score-roi-by-calibration)) or split-halves (see Figure\ \@ref(fig:E3-cal-Plot-looking-score-halves-by-calibration).

```{r E3-cal-Plot-looking-score-halves-by-calibration, echo=FALSE, warning=FALSE, fig.cap = 'Calibration scores plotted against looking scores using the left-vs.-right halves coding method (proportion of gaze samples to the half of the screen containing the new image out of all gaze samples).', out.width="50%", fig.align="center"}
ggplot(merged.eyetracking, aes(x=mean_percent_in_roi, y=looking.score.M.half))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Looking Score halves")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())

```

```{r E3-cal-Plot-looking-score-roi-by-calibration, echo=FALSE, warning=FALSE, fig.cap = 'Calibration scores plotted against looking scores using the ROI coding method (proportion of gaze samples to the new image out of all gaze samples).', out.width="50%", fig.align="center"}
ggplot(merged.eyetracking, aes(x=mean_percent_in_roi, y=looking.score.M.roi))+
  geom_point()+
  geom_smooth(method="lm", formula = y ~ x)+
  labs(x="Calibration Score", y="Looking Score (ROIs)") +
  theme_bw()+
  theme(panel.grid.minor = element_blank())

```

```{r E3-cal Prepare Day 2 data, include=FALSE}
day2.test.data <- data.day.2 %>%
  filter(task %in% c("recognition", "confidence")) %>%
  select(subject_id, image, old_or_new, correct, rt, response, task) %>%
  pivot_wider(names_from=task, values_from=c(correct, response, rt)) %>%
  select(-correct_confidence) %>%
  mutate(response_confidence = as.numeric(response_confidence))
```

```{r E3-cal Calculate Looking Score Correlations, include=FALSE}
looking.plus.behavioral <- left_join(looking.score.data, day2.test.data)

looking.correlations.subjects <- looking.plus.behavioral %>%
  filter(correct_recognition == TRUE) %>%
  filter(rt_recognition <= 10000) %>%
  group_by(subject_id) %>%
  summarize(pearson.r.conf.roi = cor(looking.score.roi, response_confidence),
            pearson.r.rt.roi = cor(looking.score.roi, rt_recognition),
            pearson.r.conf.half = cor(looking.score.half, response_confidence),
            pearson.r.rt.half = cor(looking.score.half, rt_recognition)
            ) %>%
  mutate(r.to.z.conf.roi = .5*(log(1+pearson.r.conf.roi) - log(1 - pearson.r.conf.roi)),
         r.to.z.rt.roi = .5*(log(1+pearson.r.rt.roi) - log(1 - pearson.r.rt.roi)),
         r.to.z.conf.half = .5*(log(1+pearson.r.conf.half) - log(1 - pearson.r.conf.half)),
         r.to.z.rt.half = .5*(log(1+pearson.r.rt.half) - log(1 - pearson.r.rt.half)))

looking.correlations.subjects.tidy <- looking.correlations.subjects %>%
  select(subject_id, r.to.z.conf.half, r.to.z.conf.roi, r.to.z.rt.roi, r.to.z.rt.half) %>%
  pivot_longer(2:5, names_to = "measure", values_to = "z") %>%
  mutate(behavioral_measure = case_when(
    measure %in% c("r.to.z.conf.half", "r.to.z.conf.roi") ~ "confidence",
    measure %in% c("r.to.z.rt.half", "r.to.z.rt.roi") ~ "rt",
  )) %>%
  mutate(looking_measure = case_when(
    measure %in% c("r.to.z.conf.roi", "r.to.z.rt.roi") ~ "roi",
    measure %in% c("r.to.z.rt.half", "r.to.z.conf.half") ~ "half",
  )) %>%
  select(subject_id, behavioral_measure, looking_measure, z) %>%
  left_join(subject.calibration.score, by=c("subject_id"))
```

Similarly, there was no correlation between calibration scores with the correlation between day 2 memory performance and day 1 looking for either kind of behavioral and looking measures (see Figure\ \@ref(fig:E3-cal-Plot-correlations-with-calibration-score)).

```{r E3-cal-Plot-correlations-with-calibration-score, echo=FALSE, warning=FALSE, fig.cap='The relation between calibration scores and experimental effect size, defined as the correlation between Day 1 looking scores and Day 2 memory performance. Looking scores were either coded using the halves method (left panels) or ROI method (right panels); memory performance was measured using confidence ratings (top panels) or reaction time for correct recognition judgments (bottom panels).', out.width="50%", fig.align="center"}
# Set up facet label names for behavioral_measure variable
behav.labs <- c("Confidence", "Reaction Time")
names(behav.labs) <- c("confidence", "rt")

# Set up facet label names for looking_measure variable
look.labs <- c("Halves Method", "ROI Method")
names(look.labs) <- c("half", "roi")

ggplot(looking.correlations.subjects.tidy, aes(x=mean_percent_in_roi, y=z))+
  geom_point()+
  geom_smooth(method="lm", formula=y~x)+
  facet_grid(behavioral_measure ~ looking_measure,
             labeller = labeller(behavioral_measure = behav.labs, looking_measure =
                                   look.labs))+
  labs(x="Calibration Score", y="Correlation (Looking Score, Memory)") +
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

```{r aggro exclusions}
#Identify Ps with calibration scores under 50%
validation.threshold.aggro <- 50

e3.bad.subjects.validation.aggro <- subject.calibration.score %>%
  filter(mean_percent_in_roi <= validation.threshold.aggro) %>%
  pull(subject_id) %>%
  as.character()

n.e3.bad.subjects.validation.aggro <- length(e3.bad.subjects.validation.aggro)
#26. Yikes.

#Day 1: Looking scores
t.test.data1.aggro <- t.test.data1 %>%
  filter(!subject_id %in% e3.bad.subjects.validation.aggro)

t.test.day1.aggro <- t.test(t.test.data1.aggro$M.looking.score, mu = 0.5)

#Day 2: Looking scores and Confidence
looking.confidence.correlations.subjects.filtered.aggro <- looking.confidence.correlations.subjects.filtered %>%
  filter(!subject_id %in% e3.bad.subjects.validation.aggro)

looking.confidence.correlation.t.test.aggro <- t.test(looking.confidence.correlations.subjects.filtered.aggro$r.to.z)

#Day 2: Looking scores and RT
looking.rt.correlations.subjects.filtered.aggro <- looking.rt.correlations.subjects.filtered %>%
  filter(!subject_id %in% e3.bad.subjects.validation.aggro)

looking.rt.correlation.t.test.aggro <-t.test(looking.rt.correlations.subjects.filtered.aggro$r.to.z)
```

#### Re-analysis After Exclusions
As is clear from the preceding figures, there was a large number of participants (N = `r e3.bad.subjects.validation.aggro`) that had calibration scores under 50%. When we re-analyzed the subset that remained after those participants were excluded (N = `r 45-n.e3.bad.subjects.validation.aggro`), key results were aligned with the main analyses. The looking score result was upheld such that participants looked more at the new image on Day 1 (mean: `r mean(t.test.data1$M.looking.score)` (*SD* = `r sd(t.test.data1$M.looking.score)`) (*t*(`r t.test.day1$parameter`) = `r t.test.day1$statistic`, *p* = `r t.test.day1$p.value`)), but looking scores remained unrelated to Day 2 memory outcomes (confidence: t(`r looking.confidence.correlation.t.test.aggro$parameter`) = `r looking.confidence.correlation.t.test.aggro$statistic`, p = `r looking.confidence.correlation.t.test.aggro$p.value`; RT: t(`r  looking.rt.correlation.t.test.aggro$parameter`) = `r looking.rt.correlation.t.test.aggro$statistic`, p = `r looking.rt.correlation.t.test.aggro$p.value`).