### 2024-11-03 group-e/E5-analysis-calibration.Rmd

---
title: "Experiment 5: Calibration Analysis"
output: html_notebook
---

```{r E5-cal load libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(jsonlite)
library(ggplot2)
```

```{r E5-cal load all data, include=FALSE, warning=FALSE, echo=FALSE}
data.files <- list.files('data', full.names = TRUE)
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  data.table$response <- as.character(data.table$response)
  return(data.table)
})
all.data <- bind_rows(data.tables)

# Excluding Subjects
# all.data <- all.data %>% filter(subject_id!="603ea6a51c38def3e4140ba6" & subject_id!="5ce1372c55fe0d001869c8d8" & subject_id!="5f47d860b56e68220cb86be1" & subject_id!="607eaf87ac02229ef4fa8f4a" & subject_id!="5d30857eb2f0ad00163470f3" & subject_id!="5e272684f205790453c2c4b8" & subject_id!="5f5b5881b17fa72756b8ff2a" & subject_id!="5fae5291ec402509e2885604")
# E5.excluded.n = 8

#There's an issue with Subject 603ea6a51c38def3e4140ba6 (roundness condition): their webgazer_data is a list of length 0 instead of a dataframe like the others.
all.data <- all.data %>% filter(subject_id!="603ea6a51c38def3e4140ba6")

# Code condition information
all.data <- all.data %>%
  mutate(task = if_else(study_id == '60748d36577623e2677865c6', "Attractiveness", "Roundness"))

# Convert data to long format for gaze measurements
task.data <- all.data %>%
  filter(response=="a" | response=="d") %>%
  select(subject_id, task, time_elapsed, trial_index, response, rt, webgazer_data, webgazer_targets) %>%
  unnest(webgazer_data) %>%
  tidyr::unpack(webgazer_targets) %>%
  tidyr::unpack(c('#face1', '#face2'), names_sep = ".")

# restrict to time points 1 sd before the decision
rt.summary <- all.data %>%
  filter(response=="a" | response=="d") %>%
  select(subject_id, study_id, trial_index, response, rt) %>%
  summarize(M=mean(rt), SD=sd(rt))
task.data <- task.data %>%
  mutate(time_until_decision = rt - t) %>%
  filter(time_until_decision <= rt.summary$SD)

# Bin in 50 ms bins (not 30 as in the original study, because we are using webcams)
task.data <- task.data %>%
  mutate(time.window = floor(time_until_decision/50)*50+25) #make sure bin is centered

# define ROIs
in.box <- function(x, y, left, right, top, bottom, padding){
  is.in.the.box <- x >= left - padding & x <= right + padding & y >= top - padding & y <= bottom + padding
  return(is.in.the.box)
}

#analyze with respect to ROIs
task.data <- task.data %>%
  mutate(in.left.box = in.box(x,y,`#face1.left`,`#face1.right`,`#face1.top`,`#face1.bottom`,0)) %>%
  mutate(in.right.box = in.box(x,y,`#face2.left`,`#face2.right`,`#face2.top`,`#face2.bottom`,0)) %>%
  mutate(in.target.box = if_else(response == 'a', in.left.box, in.right.box)) %>%
  mutate(in.other.box = if_else(response == 'a', in.right.box, in.left.box))

summary.gaze.view <- task.data %>%
  filter(in.target.box == TRUE | in.other.box == TRUE) %>%
  group_by(subject_id, task) %>%
  summarize(prop.in.target.subject = mean(in.target.box))

subject.calibration.score <- all.data %>%
  filter(trial_type == "webgazer-validate") %>%
  select(subject_id, trial_index, phase,  percent_in_roi, average_offset) %>% 
  unnest(percent_in_roi) %>%
  group_by(subject_id, trial_index, phase) %>% 
  summarize(mean_percent_in_roi = mean(percent_in_roi)) %>% 
  group_by(subject_id) %>% 
  mutate(calib_num = row_number()) %>%
  group_by(subject_id) %>%
  filter(calib_num == max(calib_num)) %>%
  select(subject_id, mean_percent_in_roi,calib_num)

scores <- subject.calibration.score %>%
  merge(summary.gaze.view)

scorescorrs.A <- cor.test(scores$mean_percent_in_roi[scores$task=="Attractiveness"], scores$prop.in.target.subject[scores$task=="Attractiveness"])
scorescorrs.R <- cor.test(scores$mean_percent_in_roi[scores$task=="Roundness"], scores$prop.in.target.subject[scores$task=="Roundness"])

```

As in the previous experiments, calibration score was defined as the average proportion of samples within 200 pixels of the validation point during the final validation phase before the eye tracking is performed. <!--The distribution across participants is shown in Fig. \@ref(fig:E5cal). 
(ref:E5cal-caption) Histogram of calibration success in Exp. 5. (AJ: I cut this to be more similar to the other expts)-->Where participants required more than one calibration (N=14), only the final calibration was considered.

```{r E5cal, fig.show="hold", out.width="45%", echo=FALSE,eval=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5cal-caption)", , out.width="50%", fig.align="center"}
ggplot(subject.calibration.score, aes(x=mean_percent_in_roi)) +
  stat_bin(binwidth = 10) +
  labs(x="Calibration Accuracy\n(% of samples within 200px of calibration point)") +
  theme_classic()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

```

To determine whether calibration accuracy influenced our key effects, we calculated the proportion of samples during the task in which the participant was fixating the face they ultimately chose. Calibration accuracy significantly correlated with fixations in the attractiveness condition (r = `r round(scorescorrs.A$estimate, 2)` [`r round(scorescorrs.A$conf.int, 2)[1]`, `r round(scorescorrs.A$conf.int, 2)[2]`], p = `r round(scorescorrs.A$p.value, 2)`) but not the roundness condition (r = `r round(scorescorrs.R$estimate, 2)` [`r round(scorescorrs.R$conf.int, 2)[1]`, `r round(scorescorrs.R$conf.int, 2)[2]`], p = `r round(scorescorrs.R$p.value, 2)`). Inspection of Fig. \@ref(fig:E5calcorr) reveals that this correlation is due to a handful of participants with calibration values below 50%.

(ref:E5calcorr-caption) Correlation between calibration accuracy (x-axis) and proportion of samples fixating target (y-axis) in Exp. 5.

```{r E5calcorr, fig.show="hold", out.width="80%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5calcorr-caption)", out.width="50%", fig.align="center"}
ggplot(scores, aes(x=mean_percent_in_roi, y=prop.in.target.subject, group=task, color=task))+
  geom_point() + geom_smooth(method="lm") +
  labs(x="Calibration Accuracy", y="Proportion Fixating Chosen Face") +
  #theme_classic() +
  theme_bw() +
  theme(#panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        #axis.title.y = element_blank()
        ) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )
```

```{r reanalyzedat, echo=FALSE, warning=FALSE, include=FALSE}

#re-aggregate data, removing subjects with calibration scores =< 50%.
rescore <- task.data %>%
  filter(in.target.box == TRUE | in.other.box == TRUE) %>%
  group_by(subject_id, time_until_decision, task) %>%
  summarize(prop.in.target.subject = mean(in.target.box)) %>%
  merge(subject.calibration.score) %>%
  filter(mean_percent_in_roi > 50) %>%
  mutate(time.window = floor(time_until_decision/50)*50+25) %>% #make sure bin is centered
  group_by(time.window, task) %>%
  summarize(prop.in.target = mean(prop.in.target.subject), se = sd(prop.in.target.subject)/sqrt(n()))

KS <- ks.test(rescore %>% filter(task == "Attractiveness") %>% pull(prop.in.target),
         rescore %>% filter(task == "Roundness") %>% pull(prop.in.target))

# Fitting a sigmoid as described in original

sigmoid<-function(x, a, b, c, d){
  return(d + (a - d) / (1 + (x / c)^b))
}

error <- function(x,y, pars){
  a <- pars[1]
  b <- pars[2]
  c <- pars[3]
  d <- pars[4]
  sum(sapply(1:length(x), function(i){
    return(y[i] - sigmoid(x[i], a, b, c, d))
  })^2)
}

fitsig<-function(x, y){
  pars <- c(0,0,200,0)
  return(optim(pars, fn=function(pars){error(x,y,pars)}, method="Nelder-Mead"))
}

fit.A <- fitsig(rescore$time.window[rescore$task=="Attractiveness"], rescore$prop.in.target[rescore$task=="Attractiveness"])
fit.R <- fitsig(rescore$time.window[rescore$task=="Roundness"], rescore$prop.in.target[rescore$task=="Roundness"])

fit.A.dat<-data.frame(x=rescore$time.window[rescore$task=="Attractiveness"], y=sigmoid(rescore$time.window[rescore$task=="Attractiveness"], fit.A$par[1], fit.A$par[2], fit.A$par[3], fit.A$par[4]))

fit.R.dat<-data.frame(x=rescore$time.window[rescore$task=="Roundness"], y=sigmoid(rescore$time.window[rescore$task=="Roundness"], fit.R$par[1], fit.R$par[2], fit.R$par[3], fit.R$par[4]))

# calculate R^2
R2 <- function(fit, y){
  return(1 - fit$value / sum((y - mean(y))^2))
}
R2.A<-R2(fit.A, rescore$prop.in.target[rescore$task=="Attractiveness"])
R2.R<-R2(fit.R, rescore$prop.in.target[rescore$task=="Roundness"])

plot.A<-ggplot(rescore[rescore$task=="Attractiveness",], aes(x=time.window, y=prop.in.target)) +
  geom_point(color="red") + 
  geom_line(data=fit.A.dat, aes(x=x, y=y), color="red", lwd=1.25, linetype = "solid") + 
  scale_x_reverse() + theme_classic() + 
  labs(x="Time until decision (ms)", y="Proportion of gaze in target", color="Task", fill="Task") + ylim(.4, 1) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

plot.R<-ggplot(rescore[rescore$task=="Roundness",], aes(x=time.window, y=prop.in.target))+
  geom_point(color="turquoise4") + 
  geom_line(data=fit.A.dat, aes(x=x, y=y), color="red", lwd=1.25, linetype = "dashed") + 
  geom_line(data=fit.R.dat, aes(x=x, y=y), color="turquoise4", lwd=1.25, linetype = "solid") + 
  scale_x_reverse() + theme_classic() + 
  labs(x="Time until decision (ms)", y="Proportion of gaze in target", color="Task", fill="Task") + ylim(.4, 1) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

```
#### Re-analysis After Exclusions
As in the previous experiments, we re-analyzed the data, removing the participants whose calibration accuracy was not greater than 50%. This slightly improved the fits of the sigmoids (Attractiveness: R^2^ = `r round(R2.A,2)`; Roundness: R^2^ = `r round(R2.R,2)`). However, the difference between sigmoids remained non-significant using the Kolmogorov-Smirnov test (D = `r round(KS$statistic, 2)`, p = `r round(KS$p.value, 2)`). Descriptively, the results do not look substantially different (Fig. \@ref(fig:groupEMainrev)).

(ref:groupEMainrev-caption) Revised results for Exp. 5 after removing low-calibration accuracy participants. *Left*: Eyegaze during attractiveness judgments, along with the best-fitting sigmoid. *Right*: Eyegze during roundness judgments, along with best-fitting sigmoid (best-fitting sigmoid for attractiveness is re-plotted with a dashed line for comparison).

```{r groupEMainrev, fig.show="hold", out.width="45%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:groupEMainrev-caption)"}
plot.A
plot.R
```