### 2024-11-17 group-b/E2-calibration-figure.Rmd
Added list of figure titles and captions; edited some captions.

---
title: "Experiment 2: Calibration Figure"
output: html_notebook
---

```{r include=FALSE}
library(ggplot2)
library(dplyr)
```

The task began with a 9-point eye-tracker calibration
and validation (Figure \@ref(fig:E2-calibration-figure)).

```{r include=FALSE}
# calib: [25,25],[25,75],[50,50],[75,25],[75,75]
calibration_points <- matrix(c(
  c(25,25),
  c(25,75),
  c(50,50),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)
# valid: c(25, 25, 50, 75, 75, 25, 75, 50, 25, 75)
#      [,1] [,2]
# [1,]   25   25
# [2,]   25   75
# [3,]   50   50
# [4,]   75   25
# [5,]   75   75
validation_points <- matrix(c(
  c(25,25),
  c(25,75),
  c(50,50),
  c(75,25),
  c(75,75)
), ncol=2, byrow=T)

colnames(calibration_points) <- c("x","y")
colnames(validation_points) <- c("x","y")

df <- as_tibble(calibration_points) %>% bind_rows(as_tibble(validation_points))
df$point <- c(rep("calibration", nrow(calibration_points)), rep("validation", nrow(validation_points)))
```

```{r E2-calibration-figure, echo=FALSE, fig.cap="Calibration and validation point locations for Experiment 2. Black points were used for calibration. Red crosses were used for checking the accuracy of the calibration. In this experiment, all the same locations were used for both calibration and validation.", fig.width=6, fig.height=4}
ggplot(df, aes(x=x, y=y, shape=point, color=point))+
  geom_point(size=3, stroke=1)+
  scale_shape_manual(values=c(16,3))+
  scale_color_manual(values=c("black", "red"))+
  coord_cartesian(xlim=c(0,100), ylim=c(100,0), expand=FALSE)+
  theme_void()+
  labs(color=NULL, shape=NULL)+
  theme(panel.border = element_rect(color="black", fill=NA),
        legend.position = c(0.995,0.01),
        legend.justification = c(1,0),
        legend.background = element_rect(color="black", fill=NA),
        legend.margin = margin(2,3,3,3))
```