### 2025-03-18 group-d/E4-analysis-calibration.Rmd

---
title: "E4-analysis-calibration"
author: "Rachel Ryskin"
output: pdf_document
---

```{r , include=FALSE}
library(papaja)
library(jsonlite)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
```

```{r }
data.files <- list.files('data/run-2', full.names = TRUE)
data.tables <- lapply(data.files, function(file){
  data.table <- fromJSON(file)
  return(data.table)
})
all.data <- bind_rows(data.tables)
```

```{r}

all.data.calib = all.data %>% 
  filter(trial_type == "webgazer-validate") %>% 
  dplyr::select(subject, trial_index,  percent_in_roi, average_offset) %>% 
  tidyr::unnest(percent_in_roi)

summary.data.calib = all.data.calib %>% 
  group_by(subject, trial_index) %>% 
  summarize(mean_percent_in_roi = mean(percent_in_roi)) %>% 
  group_by(subject) %>% 
  mutate(calib_num = row_number())
```

```{r, eval = F, out.width="50%", fig.align="center"}
ggplot(summary.data.calib)+
  geom_line(aes(x = calib_num, y = mean_percent_in_roi, color=subject))+
  theme_bw()+
  theme(legend.position = "none") 
```

```{r, eval=F, out.width="50%", fig.align="center"}
summary.data.calib.wide = summary.data.calib %>% 
  select(-trial_index) %>% 
  pivot_wider(id_cols = subject, names_from=calib_num, values_from = mean_percent_in_roi )

# correlation between initial and halfway calibration
ggplot(summary.data.calib.wide %>% filter(is.na(`3`)), aes(x = `1`, y = `2`))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()+
  theme(legend.position = "none")

# correlation between 2 successive calibration attempts
ggplot(summary.data.calib.wide %>% filter(!is.na(`3`)), aes(x = `1`, y = `2`))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()+
  theme(legend.position = "none")

# correlation between second attempt calibration and halfway
ggplot(summary.data.calib.wide %>% filter(!is.na(`3`)), aes(x = `2`, y = `3`))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()+
  theme(legend.position = "none") 
```

```{r}

calib.by.subj = summary.data.calib %>% 
  group_by(subject) %>% 
  summarize(mean_percent_in_roi = mean(mean_percent_in_roi))

eyetracking.effects.by.subj = read_rds( "output/E4_eye-tracking_data.rds") %>% 
  rename("condition" = compatibility) %>% 
  filter(time.window == "post-instrument-onset", condition != "filler" ) %>% 
  group_by(condition, time.window, subject) %>% 
  summarize(M = mean(prop.fixations.animal)) %>% 
  pivot_wider(names_from = condition, values_from = M) %>% 
  mutate(bias_effect = modifier - instrument) %>% 
  left_join(calib.by.subj, by = "subject")

```

```{r, eval = F, out.width="50%", fig.align="center"}
ggplot(eyetracking.effects.by.subj, aes(x = mean_percent_in_roi, y = modifier))+
  geom_point()+
  geom_smooth(method = "lm")
```

```{r}
write_rds(calib.by.subj,"output/calib_by_subj.rds")
```

```{r}

m_calib = broom::tidy(cor.test(eyetracking.effects.by.subj$mean_percent_in_roi, eyetracking.effects.by.subj$bias_effect))
#m_calib 
```

```{r}
m_calib_r = m_calib %>% pull(estimate )

```

```{r E4-calib-corr-plot, fig.cap= 'Calibration scores plotted against the verb bias effect (the difference in target animal fixations between modifier and instrument conditions).', out.width="50%", fig.align="center"}
ggplot(eyetracking.effects.by.subj, aes(x = mean_percent_in_roi, y = bias_effect))+
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x="Calibration Score", y="Bias Effect") +
  theme_bw()+
  theme(panel.grid.minor = element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

```
Participants' calibration quality, measured as the mean percentage of fixations that landed within 200 pixels of the calibration point, varied substantially (between `r range(calib.by.subj$mean_percent_in_roi) %>% round(digits = 2)`%). 
The quality of a participant's calibration significantly correlated with the participant's effect size (_Pearson's r_= `r m_calib_r`, _p_ < 0.05).
The difference in target animal fixation proportions between modifier and instrument conditions was higher for participants with better calibration (see Figure \ \@ref(fig:E4-calib-corr-plot)).

```{r read in orig data for plotting}
# only need this if we want original data in the figure comparing effect sizes between all data and good calibration subset

eyetracking.window1.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_verb.txt") %>% 
  mutate(subject = as.character(subj), 
         prop.fixations.animal = TA/(TA+TI+CA+CI),
         #sum_dur = TA+TI+CA+CI,
         condition = factor(cond, levels = c(0, 2, 1, 9), labels = c("instrument", "equibiased", "modifier", "filler")),
         time.window = "post-verb-onset-pre-animal-onset",
         study = "original") %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal, study)

eyetracking.window2.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_animal.txt")%>% 
  mutate(subject = as.character(subj), 
         prop.fixations.animal = TA/(TA+TI+CA+CI),
         condition = factor(cond, levels = c(0, 2, 1, 9), labels = c("instrument", "equibiased", "modifier", "filler")),
         time.window = "post-animal-onset-pre-instrument-onset",
         study = "original") %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal, study)

eyetracking.window3.summary.by.trial.orig = read_tsv("original_study_data/Experiment1_eye-tracking_instrument.txt")%>% 
  mutate(subject = as.character(subj), 
         prop.fixations.animal = TA/(TA+TI+CA+CI),
         condition = factor(cond, levels = c(0, 2, 1, 9), labels = c("instrument", "equibiased", "modifier", "filler")),
         time.window = "post-instrument-onset",
         study = "original") %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal, study)
```
```{r read web data}
eyetracking.window.summary.by.trial = read_rds("output/E4_eye-tracking_data.rds")
```

```{r}

eyetracking.window.summary.by.trial.all = eyetracking.window.summary.by.trial %>% 
  rename("condition" = compatibility) %>% 
  select(subject, trialID, condition, time.window, prop.fixations.animal) %>% 
  mutate(study = "web") %>% 
  left_join(calib.by.subj, by = "subject")

eyetracking.window.summary.by.trial.good.calib = eyetracking.window.summary.by.trial.all %>% 
 filter(mean_percent_in_roi >= 50) %>% 
  mutate(study = "web (calib. > 50%)")

eyetracking.window.summary.by.subj.three = bind_rows(eyetracking.window1.summary.by.trial.orig, eyetracking.window2.summary.by.trial.orig, eyetracking.window3.summary.by.trial.orig, eyetracking.window.summary.by.trial.good.calib, eyetracking.window.summary.by.trial.all) %>% 
   filter(time.window != "end", time.window != "pre-verb-onset", condition != "filler" ) %>% 
  group_by(study, condition, time.window, subject) %>% 
  summarize(mean_prop = mean(prop.fixations.animal, na.rm = T)) %>% 
  mutate(time.window = factor(time.window, 
                            levels = c("post-verb-onset-pre-animal-onset",
                                       "post-animal-onset-pre-instrument-onset",
                                       "post-instrument-onset"),
                            labels = c('Verb to Anim. \n("Rub the")',
                                       'Anim. to Instr.\n("frog with the")',
                                       'Instr.+1.5s\n(feather...")')),
         condition  = factor(condition, 
                             levels = c("modifier", "equibiased", "instrument"), 
                             labels = c("Modifier", "Equi-biased", "Instrument"))
         ) %>% 
  ungroup()

summary.eyetracking.window.summary.by.subj.three = eyetracking.window.summary.by.subj.three %>% 
  group_by(time.window, condition, study) %>% 
  summarize(prop_mean = mean(mean_prop),
            prop_se = sd(mean_prop)/sqrt(n())
            )

```
```{r E4-proportion-fix-by-window-both-post-exclusion, fig.cap = "Proportion of target fixations by verb bias in the full dataset collected online and for the subset of participants with calibration quality >50%. Error bars reflect standard errors over subject means.", out.width="75%", fig.align="center"}

shapes_legend <- c("original" = 1, "web" = 16, "web (calib. > 50%)" = 4)
linetypes_legend <- c("original" = "dashed", "web" = "solid", "web (calib. > 50%)" = "dotted")

ggplot()+
  # geom_errorbar(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "original"),
  #               aes(x=time.window,
  #                  ymin = prop_mean-prop_se,
  #                  ymax = prop_mean+prop_se,
  #                  color = condition),
  #               width = 0.1,
  #               position = position_dodge(width = .5))+
  geom_errorbar(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "web"),
                aes(x=time.window,
                   ymin = prop_mean-prop_se,
                   ymax = prop_mean+prop_se,
                   color = condition),
                width = 0.1,
                position = position_dodge(width = .5))+
  geom_errorbar(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "web (calib. > 50%)"),
                aes(x=time.window,
                   ymin = prop_mean-prop_se,
                   ymax = prop_mean+prop_se,
                   color = condition),
                width = 0.1,
                position = position_dodge(width = .5))+
  # geom_line(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "original"),
  #          aes(x=as.numeric(time.window),
  #              group = condition,
  #              y = prop_mean,
  #              linetype = "original"),
  #            alpha = 0.5,
  #           color = "grey",
  #            position = position_dodge(width = 0.5))+
 geom_line(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "web"),
           aes(x=as.numeric(time.window),
               group = condition,
               y = prop_mean,
               linetype = "web"),
             alpha = 0.5,
            color = "grey",
             position = position_dodge(width = 0.5))+
  geom_line(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "web (calib. > 50%)"),
           aes(x=as.numeric(time.window),
               group = condition,
               y = prop_mean,
               linetype = "web (calib. > 50%)"),
             alpha = 0.5,
            color = "grey",
             position = position_dodge(width = 0.5))+
  # geom_point(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "original"),
  #           aes(x=time.window, 
  #               color=condition, 
  #               y = prop_mean,
  #               shape = "original"),
  #           size = 3, 
  #           position = position_dodge(width = 0.5))+
  geom_point(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "web"),
            aes(x=time.window, 
                color=condition, 
                y = prop_mean,
                shape = "web"),
            size = 3, 
            position = position_dodge(width = 0.5))+
  geom_point(data = summary.eyetracking.window.summary.by.subj.three %>% filter(study == "web (calib. > 50%)"),
            aes(x=time.window, 
                color=condition, 
                y = prop_mean,
                shape = "web (calib. > 50%)"),
            size = 3, 
            position = position_dodge(width = 0.5))+
   scale_color_brewer(palette = "Set1")+
   scale_shape_manual(values = shapes_legend)+
   scale_linetype_manual(values = linetypes_legend)+
  # coord_cartesian(ylim = c(0,.8))+
   theme_classic()+
   labs(y = "Proportion of fixations", x = NULL, color = "Condition", shape = "Study", linetype = "Study") +     
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=13),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

```

```{r}
eyetracking.window.3 = eyetracking.window.summary.by.trial %>% 
  filter(time.window == "post-instrument-onset", compatibility != "filler" ) %>% 
  mutate(condition = factor(compatibility, levels = c('instrument', 'equibiased', 'modifier'))) %>% 
  left_join(calib.by.subj, by = "subject")
```

```{r }
# Add orthogonal contrasts to model
contrasts(eyetracking.window.3$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))
```

```{r , eval = F}

model.time.window.3 <- lmer(prop.fixations.animal ~ condition + (1 +condition | subject) + (1 | trialID), data=eyetracking.window.3,
                            control = lmerControl(optimizer = "bobyqa",
                                                  optCtrl = list(maxfun = 2e6)))
summary(model.time.window.3)
```

```{r}
eyetracking.window.3.good.calib = eyetracking.window.3 %>% 
  filter(mean_percent_in_roi >= 50)
```

```{r }
# Add orthogonal contrasts to model
contrasts(eyetracking.window.3.good.calib$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))
```

```{r }

m.good.calib <- lmer(prop.fixations.animal ~ condition + (1  | subject) + (1 | trialID), data=eyetracking.window.3.good.calib,
                            control = lmerControl(optimizer = "bobyqa",
                                                  optCtrl = list(maxfun = 2e6)))
#summary(m.good.calib)

m_calib_good = broom.mixed::tidy(m.good.calib)

m_calib_good_c1 = m_calib_good %>% filter(term == "condition1")
m_calib_good_c2 = m_calib_good %>% filter(term == "condition2")

```

#### Re-analysis After Exclusions
A subset of 35 participants had calibration quality >50%. Figure \@ref(fig:E4-proportion-fix-by-window-both-post-exclusion) shows proportions of fixations to the target animal for this subset alongside the full dataset. Though the overall proportions of fixations to the target animal in this subset were higher than in the full dataset, they were still much lower than in the original study. Replicating the linear mixed-effects analysis (in the post-instrument onset time window only) on this subset of participants suggests that the effect of verb bias condition was larger in this subset than in the full dataset. Participants' preference to the target animal relative to the target instrument in the modifier-biased condition and the equi-biased conditions was greater than in the instrument-biased condition (_b_ =  `r m_calib_good_c1 %>% pull(estimate) %>% round(digits = 2)`, _SE_ =  `r m_calib_good_c1 %>% pull(std.error) %>% round(digits = 2)`, _p_ <  0.001), but the difference betwen the modifier biased condition and the equi-biased condition was not significant (_b_ =  `r m_calib_good_c2 %>% pull(estimate) %>% round(digits = 2)`, _SE_ =  `r m_calib_good_c2 %>% pull(std.error) %>% round(digits = 2)`, _p_ =  `r m_calib_good_c2 %>% pull(p.value) %>% round(digits = 2)`).

```{r}
eyetracking.window.3.great.calib = eyetracking.window.3 %>% 
  filter(mean_percent_in_roi >= 75)
```

```{r }
# Add orthogonal contrasts to model
contrasts(eyetracking.window.3.great.calib$condition) <- cbind(c(-2/3, 1/3, 1/3), c(0, -1/2, 1/2))
```

```{r }

m.great.calib <- lmer(prop.fixations.animal ~ condition + (1  | subject) + (1 | trialID), data=eyetracking.window.3.great.calib,
                            control = lmerControl(optimizer = "bobyqa",
                                                  optCtrl = list(maxfun = 2e6)))
#summary(m.great.calib)
m_calib_great = broom.mixed::tidy(m.great.calib)

m_calib_great_c1 = m_calib_great %>% filter(term == "condition1")
m_calib_great_c2 = m_calib_great %>% filter(term == "condition2")
```

<!--AJ: Cutting this for consistency with the other Expts
Replicating the linear mixed-effects analysis (in the post-instrument onset time window only) on a subset of 19 participants with calibration quality >75% suggests that the effect of verb bias condition was larger in this subset than in the full dataset. 
Participants looked more at the target animal in the modifier-biased condition and the equi-biased conditions relative to the instrument-biased condition ( _b_ =  `r m_calib_great_c1 %>% pull(estimate) %>% round(digits = 2)`, _SE_ =  `r m_calib_great_c1 %>% pull(std.error) %>% round(digits = 2)`, _p_ <  0.001) but not significantly so in the modifier biased condition relative to the equi-biased condition ( _b_ =  `r m_calib_great_c2 %>% pull(estimate) %>% round(digits = 2)`, _SE_ =  `r m_calib_great_c2 %>% pull(std.error) %>% round(digits = 2)`, _p_ =  `r m_calib_great_c2 %>% pull(p.value) %>% round(digits = 2)`). 
-->